---
title: "Website Analytics Analysis (Plausible)"
format: html
editor: source
execute:
  echo: false
---

```{r extract_and_transform_data}
#| include: false
#| eval: false
# Set to eval: true to extract and transform data before rendering
# This chunk extracts data from Plausible API and then transforms it

library(here)

# Plausible credentials
api_key <- ""  # Or use: Sys.getenv("PLAUSIBLE_API_KEY")
site_id <- "ds4owd-002.github.io/website"

if (api_key != "" && site_id != "") {
  # Step 1: Extract data using bash script
  extraction_script <- here("scripts", "01_extract_plausible_data.sh")
  output_json <- here("data", "plausible_raw.json")

  extract_result <- system2(
    extraction_script,
    args = c(shQuote(api_key), shQuote(site_id), output_json),
    stdout = TRUE,
    stderr = TRUE
  )

  if (!is.null(attr(extract_result, "status")) && attr(extract_result, "status") != 0) {
    cat("Error extracting data:\n")
    cat(paste(extract_result, collapse = "\n"), "\n")
    stop("Failed to extract Plausible data")
  }

  # Step 2: Transform the extracted data
  source(here("scripts", "02_transform_plausible_data.R"))

  cat("Data extraction and transformation complete!\n")
} else {
  cat("Note: API key or site ID not provided.\n")
  cat("Skipping data extraction. Using existing processed data.\n")
}
```

```{r setup}
#| include: false
library(tidyverse)
library(lubridate)
library(scales)
library(ggthemes)
library(here)

# Load data
load(here("data", "plausible_data.rda"))

# Set theme
theme_set(theme_minimal())
```

## Overview

This analysis covers website analytics for **`r plausible_metadata$site_id`**, spanning **`r plausible_metadata$total_days` days** from **`r format(plausible_metadata$date_range_start, "%B %d, %Y")`** to **`r format(plausible_metadata$date_range_end, "%B %d, %Y")`**.

*Data extracted: `r format(plausible_metadata$extracted_at, "%B %d, %Y at %H:%M UTC")`*

**Overall Statistics:**

- Total unique visitors: **`r comma(summary_stats$total_visitors)`**
- Total visits: **`r comma(summary_stats$total_visits)`**
- Total pageviews: **`r comma(summary_stats$total_pageviews)`**
- Average bounce rate: `r round(summary_stats$avg_bounce_rate, 1)`%
- Average visit duration: `r round(summary_stats$avg_visit_duration)` seconds

## Traffic Trends Over Time

### Daily Visitors

```{r daily_visitors}
#| fig-width: 10
#| fig-height: 6

timeseries_data |>
  ggplot(aes(x = date, y = visitors)) +
  geom_line(color = colorblind_pal()(8)[1], linewidth = 1) +
  geom_point(color = colorblind_pal()(8)[1], size = 1.5, alpha = 0.5) +
  labs(title = "Daily Unique Visitors",
       x = "Date",
       y = "Unique Visitors") +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Daily Pageviews

```{r daily_pageviews}
#| fig-width: 10
#| fig-height: 6

timeseries_data |>
  ggplot(aes(x = date, y = pageviews)) +
  geom_line(color = colorblind_pal()(8)[2], linewidth = 1) +
  geom_point(color = colorblind_pal()(8)[2], size = 1.5, alpha = 0.5) +
  labs(title = "Daily Pageviews",
       x = "Date",
       y = "Pageviews") +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Multiple Metrics Over Time

```{r multi_metrics}
#| fig-width: 10
#| fig-height: 6

# Prepare data
multi_metrics_data <- timeseries_data |>
  select(date, visitors, visits, pageviews) |>
  pivot_longer(cols = -date, names_to = "metric", values_to = "count")

# Create plot
multi_metrics_data |>
  ggplot(aes(x = date, y = count, color = metric)) +
  geom_line(linewidth = 1) +
  labs(title = "Traffic Metrics Over Time",
       x = "Date",
       y = "Count",
       color = "Metric") +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  scale_color_colorblind(labels = c("Pageviews", "Visitors", "Visits")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Engagement Metrics Over Time

```{r engagement_over_time}
#| fig-width: 10
#| fig-height: 6

timeseries_data |>
  ggplot(aes(x = date, y = bounce_rate)) +
  geom_line(color = colorblind_pal()(8)[4], linewidth = 1) +
  labs(title = "Bounce Rate Over Time",
       x = "Date",
       y = "Bounce Rate (%)") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r visit_duration_over_time}
#| fig-width: 10
#| fig-height: 6

# Prepare data
visit_duration_data <- timeseries_data |>
  mutate(visit_duration_min = visit_duration / 60)

# Create plot
visit_duration_data |>
  ggplot(aes(x = date, y = visit_duration_min)) +
  geom_line(color = colorblind_pal()(8)[6], linewidth = 1) +
  labs(title = "Average Visit Duration Over Time",
       x = "Date",
       y = "Visit Duration (minutes)") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Monthly Analysis

```{r monthly_trends}
#| fig-width: 10
#| fig-height: 6

# Prepare data
monthly_trends_data <- timeseries_data |>
  mutate(month = floor_date(date, "month")) |>
  group_by(month) |>
  summarise(
    visitors = sum(visitors, na.rm = TRUE),
    visits = sum(visits, na.rm = TRUE),
    pageviews = sum(pageviews, na.rm = TRUE),
    avg_bounce_rate = mean(bounce_rate, na.rm = TRUE),
    avg_visit_duration = mean(visit_duration, na.rm = TRUE)
  )

# Create plot
monthly_trends_data |>
  ggplot(aes(x = month, y = visitors)) +
  geom_col(fill = colorblind_pal()(8)[2]) +
  labs(title = "Monthly Unique Visitors",
       x = "Month",
       y = "Unique Visitors") +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_breaks = "1 month", date_labels = "%b %Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r monthly_table}
#| echo: false

timeseries_data |>
  mutate(month = floor_date(date, "month")) |>
  group_by(month) |>
  summarise(
    Visitors = sum(visitors, na.rm = TRUE),
    Visits = sum(visits, na.rm = TRUE),
    Pageviews = sum(pageviews, na.rm = TRUE),
    `Bounce Rate (%)` = round(mean(bounce_rate, na.rm = TRUE), 1),
    `Avg Duration (s)` = round(mean(visit_duration, na.rm = TRUE), 0)
  ) |>
  arrange(desc(month)) |>
  mutate(month = format(month, "%b %Y")) |>
  rename(Month = month) |>
  knitr::kable(caption = "Monthly Statistics")
```

## Traffic Sources

```{r traffic_sources}
#| fig-width: 10
#| fig-height: 6

# Prepare data
traffic_sources_plot_data <- sources_data |>
  head(10) |>
  mutate(
    source = fct_reorder(source, visitors),
    pct = visitors / sum(sources_data$visitors) * 100
  )

# Recycle colorblind palette colors for all sources
n_sources <- nrow(traffic_sources_plot_data)
colors <- rep(colorblind_pal()(8), length.out = n_sources)

# Create plot
traffic_sources_plot_data |>
  ggplot(aes(x = visitors, y = source, fill = source)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(comma(visitors), " (", round(pct, 1), "%)")),
            hjust = -0.1, size = 4) +
  labs(title = "Top 10 Visitors by Traffic Source",
       x = "Unique Visitors",
       y = "Source") +
  scale_x_continuous(labels = comma) +
  scale_fill_manual(values = colors) +
  expand_limits(x = max(traffic_sources_plot_data$visitors) * 1.2)
```

```{r sources_metrics}
#| fig-width: 10
#| fig-height: 6

# Prepare data
sources_metrics_plot_data <- sources_data |>
  head(10) |>
  mutate(source = fct_reorder(source, bounce_rate))

# Recycle colorblind palette colors for all sources
n_sources <- nrow(sources_metrics_plot_data)
colors <- rep(colorblind_pal()(8), length.out = n_sources)

# Create plot
sources_metrics_plot_data |>
  ggplot(aes(x = bounce_rate, y = source, fill = source)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(round(bounce_rate, 1), "%")),
            hjust = -0.2, size = 4) +
  labs(title = "Bounce Rate by Traffic Source (Top 10)",
       x = "Bounce Rate (%)",
       y = "Source") +
  scale_fill_manual(values = colors) +
  expand_limits(x = max(sources_metrics_plot_data$bounce_rate) * 1.15)
```

```{r sources_table}
#| echo: false
#| include: false

sources_data |>
  mutate(
    `Share (%)` = round(visitors / sum(visitors) * 100, 1)
  ) |>
  select(
    Source = source,
    Visitors = visitors,
    `Share (%)`,
    Visits = visits,
    Pageviews = pageviews,
    `Bounce Rate (%)` = bounce_rate,
    `Avg Duration (s)` = visit_duration
  ) |>
  mutate(`Bounce Rate (%)` = round(`Bounce Rate (%)`, 1)) |>
  knitr::kable(caption = "Traffic Source Statistics")
```

## Device Analysis

```{r devices}
#| fig-width: 8
#| fig-height: 5

# Prepare data
devices_plot_data <- devices_data |>
  mutate(
    pct = visitors / sum(visitors) * 100,
    device = fct_reorder(device, visitors)
  )

# Create plot
devices_plot_data |>
  ggplot(aes(x = visitors, y = device, fill = device)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(comma(visitors), "\n(", round(pct, 1), "%)")),
            hjust = -0.1, size = 4) +
  labs(title = "Visitors by Device Type",
       x = "Unique Visitors",
       y = "Device") +
  scale_x_continuous(labels = comma) +
  scale_fill_colorblind() +
  expand_limits(x = max(devices_plot_data$visitors) * 1.2)
```

```{r device_metrics}
#| echo: false

devices_data |>
  mutate(
    `Share (%)` = round(visitors / sum(visitors) * 100, 1)
  ) |>
  select(
    Device = device,
    Visitors = visitors,
    `Share (%)`,
    Pageviews = pageviews,
    `Bounce Rate (%)` = bounce_rate,
    `Avg Duration (s)` = visit_duration
  ) |>
  mutate(`Bounce Rate (%)` = round(`Bounce Rate (%)`, 1)) |>
  knitr::kable(caption = "Device Statistics")
```

## Browser & Operating System

```{r browsers}
#| fig-width: 10
#| fig-height: 6

# Prepare data
browsers_plot_data <- browsers_data |>
  head(5) |>
  mutate(
    browser = fct_reorder(browser, visitors),
    pct = visitors / sum(browsers_data$visitors) * 100
  )

# Recycle colorblind palette colors for all browsers
n_browsers <- nrow(browsers_plot_data)
colors <- rep(colorblind_pal()(8), length.out = n_browsers)

# Create plot
browsers_plot_data |>
  ggplot(aes(x = visitors, y = browser, fill = browser)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(comma(visitors), " (", round(pct, 1), "%)")),
            hjust = -0.1, size = 3.5) +
  labs(title = "Top 5 Browsers by Unique Visitors",
       x = "Unique Visitors",
       y = "Browser") +
  scale_x_continuous(labels = comma) +
  scale_fill_manual(values = colors) +
  expand_limits(x = max(browsers_plot_data$visitors) * 1.25)
```

```{r os}
#| fig-width: 10
#| fig-height: 6

# Prepare data
os_plot_data <- os_data |>
  head(5) |>
  mutate(
    os = fct_reorder(os, visitors),
    pct = visitors / sum(os_data$visitors) * 100
  )

# Recycle colorblind palette colors for all OS
n_os <- nrow(os_plot_data)
colors <- rep(colorblind_pal()(8), length.out = n_os)

# Create plot
os_plot_data |>
  ggplot(aes(x = visitors, y = os, fill = os)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(comma(visitors), " (", round(pct, 1), "%)")),
            hjust = -0.1, size = 3.5) +
  labs(title = "Top 5 Operating Systems by Unique Visitors",
       x = "Unique Visitors",
       y = "Operating System") +
  scale_x_continuous(labels = comma) +
  scale_fill_manual(values = colors) +
  expand_limits(x = max(os_plot_data$visitors) * 1.25)
```

## Geographic Distribution

```{r countries}
#| fig-width: 10
#| fig-height: 8

# Prepare data
countries_plot_data <- countries_data |>
  head(20) |>
  mutate(
    country = fct_reorder(country, visitors),
    pct = visitors / sum(countries_data$visitors) * 100
  )

# Recycle colorblind palette colors for all countries
n_countries <- nrow(countries_plot_data)
colors <- rep(colorblind_pal()(8), length.out = n_countries)

# Create plot
countries_plot_data |>
  ggplot(aes(x = visitors, y = country, fill = country)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(comma(visitors), " (", round(pct, 1), "%)")),
            hjust = -0.1, size = 3.5) +
  labs(title = "Top 20 Countries by Unique Visitors",
       x = "Unique Visitors",
       y = "Country") +
  scale_x_continuous(labels = comma) +
  scale_fill_manual(values = colors) +
  expand_limits(x = max(countries_plot_data$visitors) * 1.25)
```

```{r countries_table}
#| echo: false
#| include: false

countries_data |>
  head(20) |>
  mutate(
    `Share (%)` = round(visitors / sum(countries_data$visitors) * 100, 1)
  ) |>
  select(
    Country = country,
    Visitors = visitors,
    `Share (%)`,
    Visits = visits,
    Pageviews = pageviews
  ) |>
  knitr::kable(caption = "Top 20 Countries Statistics")
```

## Top Pages

```{r top_pages}
#| fig-width: 10
#| fig-height: 6

# Prepare data
top_pages_plot_data <- pages_data |>
  head(10) |>
  mutate(
    page = fct_reorder(page, visitors),
    pct = visitors / sum(visitors) * 100
  )

# Recycle colorblind palette colors for all pages
n_pages <- nrow(top_pages_plot_data)
colors <- rep(colorblind_pal()(8), length.out = n_pages)

# Create plot
top_pages_plot_data |>
  ggplot(aes(x = visitors, y = page, fill = page)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = comma(visitors)),
            hjust = -0.1, size = 3.5) +
  labs(title = "Top 10 Pages by Unique Visitors",
       x = "Unique Visitors",
       y = "Page") +
  scale_x_continuous(labels = comma) +
  scale_fill_manual(values = colors) +
  expand_limits(x = max(top_pages_plot_data$visitors) * 1.15)
```

```{r pages_engagement}
#| fig-width: 10
#| fig-height: 6

# Prepare data
pages_engagement_plot_data <- pages_data |>
  filter(!is.na(bounce_rate)) |>
  head(5) |>
  mutate(page = fct_reorder(page, bounce_rate))

# Recycle colorblind palette colors for all pages
n_pages <- nrow(pages_engagement_plot_data)
colors <- rep(colorblind_pal()(8), length.out = n_pages)

# Create plot
pages_engagement_plot_data |>
  ggplot(aes(x = bounce_rate, y = page, fill = page)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(round(bounce_rate, 1), "%")),
            hjust = -0.1, size = 3.5) +
  labs(title = "Bounce Rate by Page (Top 5)",
       x = "Bounce Rate (%)",
       y = "Page") +
  scale_fill_manual(values = colors) +
  expand_limits(x = max(pages_engagement_plot_data$bounce_rate) * 1.15)
```

```{r pages_table}
#| echo: false

pages_data |>
  head(10) |>
  mutate(
    `Share (%)` = round(visitors / sum(visitors) * 100, 1)
  ) |>
  select(
    Page = page,
    Visitors = visitors,
    `Share (%)`,
    Pageviews = pageviews,
    `Bounce Rate (%)` = bounce_rate,
    `Avg Duration (s)` = visit_duration
  ) |>
  mutate(`Bounce Rate (%)` = round(`Bounce Rate (%)`, 1)) |>
  knitr::kable(caption = "Top 10 Pages Statistics")
```

## Day of Week Analysis

```{r day_of_week}
#| fig-width: 10
#| fig-height: 6

# Prepare data
day_of_week_data <- timeseries_data |>
  mutate(
    weekday = wday(date, label = TRUE, abbr = FALSE),
    weekday = fct_relevel(weekday, "Monday", "Tuesday", "Wednesday",
                          "Thursday", "Friday", "Saturday", "Sunday")
  ) |>
  group_by(weekday) |>
  summarise(
    avg_visitors = mean(visitors, na.rm = TRUE),
    avg_pageviews = mean(pageviews, na.rm = TRUE)
  )

# Create plot
day_of_week_data |>
  ggplot(aes(x = weekday, y = avg_visitors, fill = weekday)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = comma(round(avg_visitors, 0))),
            vjust = -0.5, size = 4) +
  labs(title = "Average Daily Visitors by Day of Week",
       x = "Day of Week",
       y = "Average Visitors") +
  scale_y_continuous(labels = comma) +
  scale_fill_colorblind() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  expand_limits(y = max(day_of_week_data$avg_visitors) * 1.15)
```

## Summary Insights

```{r summary_insights_calc}
#| echo: false
#| include: false

# Calculate growth metrics
has_growth_data <- FALSE
if (nrow(timeseries_data) > 30) {
  recent_30 <- timeseries_data |>
    arrange(desc(date)) |>
    head(30) |>
    summarise(avg_visitors = mean(visitors, na.rm = TRUE))

  previous_30 <- timeseries_data |>
    arrange(desc(date)) |>
    slice(31:60) |>
    summarise(avg_visitors = mean(visitors, na.rm = TRUE))

  if (nrow(previous_30) > 0 && previous_30$avg_visitors > 0) {
    growth_rate <- ((recent_30$avg_visitors - previous_30$avg_visitors) /
                    previous_30$avg_visitors) * 100
    has_growth_data <- TRUE
  }
}

# Calculate device percentage
device_pct <- round(devices_data$visitors[1] / sum(devices_data$visitors) * 100, 1)
```

**30-Day Growth Rate:**

- Recent 30 days avg: `r round(recent_30$avg_visitors, 0)` visitors/day
- Previous 30 days avg: `r round(previous_30$avg_visitors, 0)` visitors/day
- Growth rate: `r round(growth_rate, 1)`%

**Top Performers:**

- Best traffic source: **`r sources_data$source[1]`** with `r comma(sources_data$visitors[1])` visitors
- Most popular page: **`r pages_data$page[1]`** with `r comma(pages_data$visitors[1])` visitors
- Top country: **`r countries_data$country[1]`** with `r comma(countries_data$visitors[1])` visitors
- Dominant device: **`r devices_data$device[1]`** with `r device_pct`% of visitors
