---
title: "GHE on LinkedIn"
format:
  html:
    code-fold: true
    code-summary: "Display code"
    toc: true
    toc-depth: 4
    toc-location: left
execute:
  message: false
  warning: false
---

While statistics can be fetched from LinkedIn either via their API or manually from the overview page, their scope is limited. Statistics are only available for a rolling 12-month window. This means data must be collected at regular intervals to avoid gaps.

## Data Processing Pipeline

::: {style="width: 100%; max-width: 1200px;"}
```{mermaid}
%%| fig-width: 10
flowchart TD
    raw["raw-data/"]

    subgraph script["data-prep.R"]
        p1["Content Metrics Processing"]
        p2["All Posts Processing"]
        p3["New Followers Processing"]
    end

    subgraph clean["clean-data/"]
        c1["content-overview.csv"]
        c2["content_posts-overview.csv"]
        c3["new-followers.csv"]
    end

    raw --> p1
    raw --> p2
    raw --> p3

    p1 --> c1
    p2 --> c2
    p3 --> c3

    c1 & c2 & c3 --> analysis["analysis.qmd"]

    style raw fill:#e1f5ff
    style script fill:#fff4e1
    style clean fill:#e8f5e9
```
:::

We download our analytics data manually from LinkedIn roughly every three months. On the group dashboard (admin rights required), there are three tabs: "Content", "Visitors", and "Followers". These datasets serve as the source for our metadata and are stored in `raw-data`. Due to the historical limits mentioned above, we continuously append the latest data to the existing dataset.

This README must be rendered with the following command in your terminal: `quarto render analysis.qmd --to gfm --output README.md`

## Followers

```{r, message=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggthemes)

followers  <- read_csv("clean-data/new-followers.csv")
content_posts  <- read_csv("clean-data/content_posts-overview.csv")
content_overview  <- read_csv("clean-data/content_posts-overview.csv")
```

```{r}
followers |>
  ggplot(aes(x = date, y = cumulative_followers)) +
  geom_line() +
  labs(
    x = "",
    y = "Followers\n"
  ) +
    theme_few()
```

## Content

### Impressions

```{r}
# impressions over time
content_overview %>%
  group_by(month = floor_date(date, "week")) %>%
  summarise(
    sum_value = sum(impressions, na.rm = TRUE),
    mean_value = mean(impressions, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ggplot(aes(x = month, y = sum_value)) +
  geom_smooth() +
  geom_col() +
  labs(
    x = "",
    y = "Impressions\n"
  ) +
  theme_few()
```

### Likes

```{r}
content_overview %>%
  group_by(month = floor_date(date, "week")) %>%
  summarise(
    sum_value = sum(likes, na.rm = TRUE),
    mean_value = mean(likes, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ggplot(aes(x = month, y = sum_value)) +
  geom_smooth() +
  geom_col() +
  labs(
    x = "",
    y = "Likes\n"
  ) +
  theme_few()
```

### Likes per impression

```{r}
content_overview  |> 
  group_by(month = floor_date(date, "week")) %>%
  summarise(
    sum_likes = sum(likes, na.rm = TRUE),
    sum_impressions = sum(impressions, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  mutate(likes_per_impression = sum_likes/sum_impressions) |>
  ggplot(aes(x = month, y = likes_per_impression))  +
    geom_col() +
  geom_smooth() +
  labs(
    x = "",
    y = "Likes\n"
  ) +
  theme_few()
```

### Number of posts

```{r}
mean_n_posts  <- content_overview %>%
  group_by(month = floor_date(date, "month")) %>%
  count() |> 
  ungroup() |> 
  summarise(mean_posts = mean(n)) |> 
  pull()

content_overview %>%
  group_by(month = floor_date(date, "month")) %>%
  count() |> 
  ggplot(aes(x = month, y = n)) +
  geom_col() +
  geom_hline(yintercept = mean_n_posts, color = "red", linetype = "dotted") +
  labs(
    x = "",
    y = "Number of posts per month\n"
  ) +
  theme_few()
```

### Impressions per post

```{r}
content_overview %>%
  group_by(month = floor_date(date, "week")) %>%
  summarise(n_posts = n(),
  impression = sum(impressions)) |> 
    mutate(impressions_per_post = impression/n_posts) |> 
  ggplot(aes(x = month, y = impressions_per_post)) +
  geom_smooth() +
  geom_col() +
  labs(x = "",
  y = "Impressions\n") +
  theme_few()
```

### Top 5: Likes

```{r}
# filter top 10 posts according to likes
content_posts |> 
  slice_max(likes, n = 5) |> 
  select(post_title, likes) |> 
  knitr::kable(format = "html")
```

### Top 5: Reposts

```{r}
# filter top 10 posts according to likes
content_posts |> 
  slice_max(reposts, n = 5) |> 
  select(post_title, reposts) |> 
  knitr::kable(format = "html")
```

### Top 5: Engagement Rate

```{r}
content_posts |> 
  slice_max(engagement_rate, n = 5) |> 
  select(post_title, engagement_rate) |>
  arrange(desc(engagement_rate)) |>
  knitr::kable(format = "html")
```

