---
title: "GHE on LinkedIn"
format:
  html:
    code-fold: true
    code-summary: "Display code"
    toc: true
    toc-depth: 4
    toc-location: left
execute:
  message: false
  warning: false
---

While statistics can be fetched from LinkedIn either via their API or manually from the overview page, their scope is limited. Statistics are only available for a rolling 12-month window. This means data must be collected at regular intervals to avoid gaps.

## Data Processing Pipeline

::: {style="width: 100%; max-width: 1200px;"}
```{mermaid}
%%| fig-width: 10
flowchart TD
    raw["raw-data/"]

    subgraph script["data-prep.R"]
        p1["Content Metrics Processing"]
        p2["All Posts Processing"]
        p3["New Followers Processing"]
        p4["Visitors Metrics Processing"]
    end

    subgraph clean["clean-data/"]
        c1["content-overview.csv"]
        c2["content_posts-overview.csv"]
        c3["new-followers.csv"]
        c4["visitors-overview.csv"]
    end

    raw --> p1
    raw --> p2
    raw --> p3
    raw --> p4

    p1 --> c1
    p2 --> c2
    p3 --> c3
    p4 --> c4

    c1 & c2 & c3 & c4 --> analysis["analysis.qmd"]

    style raw fill:#e1f5ff
    style script fill:#fff4e1
    style clean fill:#e8f5e9
```
:::

We download our analytics data manually from LinkedIn roughly every three months. On the group dashboard (admin rights required), there are three tabs: "Content", "Visitors", and "Followers". These datasets serve as the source for our metadata and are stored in `raw-data`. Due to the historical limits mentioned above, we continuously append the latest data to the existing dataset.

This README must be rendered with the following command in your terminal: `quarto render analysis.qmd --to gfm --output README.md`

## Followers

```{r, message=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggthemes)

followers  <- read_csv("clean-data/new-followers.csv")
content_posts  <- read_csv("clean-data/content_posts-overview.csv")
content_overview  <- read_csv("clean-data/content_posts-overview.csv")
visitors  <- read_csv("clean-data/visitors-overview.csv")
```

```{r}
followers |>
  ggplot(aes(x = date, y = cumulative_followers)) +
  geom_line() +
  labs(
    x = "",
    y = "Followers\n"
  ) +
    theme_few()
```

## Content

### Impressions

```{r}
# impressions over time
content_overview %>%
  group_by(month = floor_date(date, "week")) %>%
  summarise(
    sum_value = sum(impressions, na.rm = TRUE),
    mean_value = mean(impressions, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ggplot(aes(x = month, y = sum_value)) +
  geom_smooth() +
  geom_col() +
  labs(
    x = "",
    y = "Impressions\n"
  ) +
  theme_few()
```

### Likes

```{r}
content_overview %>%
  group_by(month = floor_date(date, "week")) %>%
  summarise(
    sum_value = sum(likes, na.rm = TRUE),
    mean_value = mean(likes, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ggplot(aes(x = month, y = sum_value)) +
  geom_smooth() +
  geom_col() +
  labs(
    x = "",
    y = "Likes\n"
  ) +
  theme_few()
```

### Likes per impression

```{r}
content_overview  |> 
  group_by(month = floor_date(date, "week")) %>%
  summarise(
    sum_likes = sum(likes, na.rm = TRUE),
    sum_impressions = sum(impressions, na.rm = TRUE),
    .groups = "drop"
  ) |> 
  mutate(likes_per_impression = sum_likes/sum_impressions) |>
  ggplot(aes(x = month, y = likes_per_impression))  +
    geom_col() +
  geom_smooth() +
  labs(
    x = "",
    y = "Likes\n"
  ) +
  theme_few()
```

### Number of posts

```{r}
mean_n_posts  <- content_overview %>%
  group_by(month = floor_date(date, "month")) %>%
  count() |> 
  ungroup() |> 
  summarise(mean_posts = mean(n)) |> 
  pull()

content_overview %>%
  group_by(month = floor_date(date, "month")) %>%
  count() |> 
  ggplot(aes(x = month, y = n)) +
  geom_col() +
  geom_hline(yintercept = mean_n_posts, color = "red", linetype = "dotted") +
  labs(
    x = "",
    y = "Number of posts per month\n"
  ) +
  theme_few()
```

### Impressions per post

```{r}
content_overview %>%
  group_by(month = floor_date(date, "week")) %>%
  summarise(n_posts = n(),
  impression = sum(impressions)) |> 
    mutate(impressions_per_post = impression/n_posts) |> 
  ggplot(aes(x = month, y = impressions_per_post)) +
  geom_smooth() +
  geom_col() +
  labs(x = "",
  y = "Impressions\n") +
  theme_few()
```

### Impressions and Followers

```{r}
followers_week  <- followers |> 
  mutate(week = floor_date(date, unit = "week", week_start = 1)) |> 
  group_by(week) |> 
  slice_max(cumulative_followers) |> 
  distinct(week, .keep_all = TRUE)  |> 
  select(week, cumulative_followers)

content_week  <- content_overview |> 
  mutate(week = floor_date(date, unit = "week", week_start = 1)) |> 
  group_by(week) |> 
  summarise(total_impressions = sum(impressions))

merge_week  <- followers_week |> 
  left_join(content_week)

merge_week  |> 
  ggplot() +
  geom_col(data = content_week, aes(x = week, y = total_impressions)) +
  geom_point(data = followers_week, aes(x = week, y = cumulative_followers), color = "red") +
  labs(x = "",
  y = "Weekly Impressions\n") +
    theme_few()

cor_data  <- merge_week |> 
  drop_na()

cor(cor_data$cumulative_followers, cor_data$total_impressions)
```

## Deep dive on high performing weeks

```{r}
impressions_deepdive  <- content_overview |> 
  mutate(week = floor_date(date, unit = "week", week_start = 1)) |> 
  group_by(week) |> 
  mutate(total_impressions = sum(impressions)) |> 
  filter(total_impressions > 7000)
```

```{r}
content_overview |> 
  ggplot(aes(x = impressions, y = likes)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_few()

cor(content_overview$impressions, content_overview$likes)
```

```{r}
content_overview_less7k <- content_overview |> 
  filter(impressions < 7000)

content_overview_less7k |> 
  ggplot(aes(x = impressions, y = likes)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_few()

cor(content_overview_less7k$impressions, content_overview_less7k$likes)
```

```{r}
paste0("Correlation Impressions / Likes: ", cor(content_overview$impressions, content_overview$likes))
paste0("Correlation Impressions / Comments: ", cor(content_overview$impressions, content_overview$comments))
paste0("Correlation Impressions / Reposts: ", cor(content_overview$impressions, content_overview$reposts))
paste0("Correlation Impressions / Clicks: ", cor(content_overview$impressions, content_overview$clicks))
paste0("Correlation Impressions / Engagement Rate: ", cor(content_overview$impressions, content_overview$engagement_rate))
```

### Top 5: Likes

```{r}
# filter top 10 posts according to likes
content_posts |> 
  slice_max(likes, n = 5) |> 
  select(post_title, likes) |> 
  knitr::kable(format = "html")
```

### Top 5: Reposts

```{r}
# filter top 10 posts according to likes
content_posts |> 
  slice_max(reposts, n = 5) |> 
  select(post_title, reposts) |> 
  knitr::kable(format = "html")
```

### Top 5: Engagement Rate

```{r}
content_posts |>
  slice_max(engagement_rate, n = 5) |>
  select(post_title, engagement_rate) |>
  arrange(desc(engagement_rate)) |>
  knitr::kable(format = "html")
```

### Click Through Rate

```{r}
content_overview |>
  slice_max(n = 5, order_by = click_through_rate_ctr) |>
  select(post_title, click_through_rate_ctr) |>
  knitr::kable(format = "html")
```

```{r}
hist(content_overview$click_through_rate_ctr, main = "Distribution of CTR", xlab = "Click Through Rate")
```

```{r}
content_overview |>
  ggplot(aes(x = impressions, y = click_through_rate_ctr)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "\nImpressions", y = "Click Through Rate\n") +
  theme_few()

paste0("Correlation: ", cor(content_overview$impressions, content_overview$click_through_rate_ctr))
round(quantile(x = content_overview$click_through_rate_ctr, probs = seq(0,1,.1)), digits = 2)
```

```{r}
click_through_no_outliers  <- content_overview |>
  filter(click_through_rate_ctr < 0.4)

click_through_no_outliers |>
  ggplot(aes(x = impressions, y = click_through_rate_ctr)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "\nImpressions", y = "Click Through Rate\n") +
  theme_few()

paste0("Correlation: ", cor(click_through_no_outliers$impressions, click_through_no_outliers$click_through_rate_ctr))
```

## Visitors

### Total page views per week

```{r}
visitors |>
  group_by(week = floor_date(date, "week")) |>
  summarise(page_views = sum(total_page_views, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = week, y = page_views)) +
  geom_col() +
  geom_smooth() +
  labs(
    x = "",
    y = "Page views\n"
  ) +
  theme_few()
```

### Total unique visitors per week

```{r}
visitors |>
  group_by(week = floor_date(date, "week")) |>
  summarise(unique_visitors = sum(total_unique_visitors, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = week, y = unique_visitors)) +
  geom_col() +
  geom_smooth() +
  labs(
    x = "",
    y = "Unique visitors\n"
  ) +
  theme_few()
```

### Page views by subpage

```{r}
visitors |>
  group_by(week = floor_date(date, "week")) |>
  summarise(
    Overview = sum(overview_page_views, na.rm = TRUE),
    Life = sum(life_page_views, na.rm = TRUE),
    Jobs = sum(jobs_page_views, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(-week, names_to = "subpage", values_to = "page_views") |>
  ggplot(aes(x = week, y = page_views, fill = subpage)) +
  geom_col() +
  labs(
    x = "",
    y = "Page views\n",
    fill = ""
  ) +
  theme_few()
```

### Desktop vs. mobile page views

```{r}
visitors |>
  group_by(week = floor_date(date, "week")) |>
  summarise(
    Desktop = sum(total_page_views_desktop, na.rm = TRUE),
    Mobile = sum(total_page_views_mobile, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_longer(-week, names_to = "device", values_to = "page_views") |>
  ggplot(aes(x = week, y = page_views, fill = device)) +
  geom_col() +
  labs(
    x = "",
    y = "Page views\n",
    fill = ""
  ) +
  theme_few()
```
