---
title: "GHE Supervision Analysis"
author: "Global Health Engineering"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
execute:
  echo: true
  warning: false
  message: false
code-fold: true
---

```{r extract_data}
#| include: false
#| eval: false
# Set to eval: true to extract and transform data before rendering
# This chunk extracts data from Google Sheets and saves it as CSV

library(here)
source(here("scripts", "01_extract_transform_supervision_data.R"))
```

```{r setup}
library(tidyverse)
library(lubridate)
library(scales)
library(knitr)
library(here)

# Load data
people <- read_csv(here("data", "people.csv"), show_col_types = FALSE)

# Derive year from end_date if year column is all NA
if (all(is.na(people$year))) {
  people <- people |>
    mutate(year = year(end_date))
}

# Set theme for all plots
theme_set(theme_minimal())
```

## Overview

This analysis covers the supervision records from the GHE research group. Names and identifiers have been removed for privacy reasons.

```{r overview_stats}
#| echo: false
#| include: false

n_total <- nrow(people)
n_degrees <- n_distinct(people$degree)
n_archived <- sum(people$status == "archived", na.rm = TRUE)
n_in_progress <- sum(people$status == "in-progress", na.rm = TRUE)
year_range <- if (any(!is.na(people$year))) range(people$year, na.rm = TRUE) else c(NA, NA)
```

**Key Statistics:**

- Total supervision records: `r n_total`
- Unique degree types: `r n_degrees`
- Year range: `r if (!is.na(year_range[1])) paste(year_range[1], "to", year_range[2]) else "N/A"`
- Archived: `r n_archived`
- In progress: `r n_in_progress`

## Supervision by Degree Type

```{r degree_distribution}
#| fig-width: 10
#| fig-height: 6

# Prepare data
degree_data <- people |>
  filter(!is.na(degree)) |>
  count(degree, sort = TRUE) |>
  mutate(
    degree = fct_reorder(degree, n),
    pct = round(n / sum(n) * 100, 1)
  )

# Create plot
degree_data |>
  ggplot(aes(x = n, y = degree, fill = degree)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(n, " (", pct, "%)")),
            hjust = -0.1, size = 4) +
  labs(title = "Distribution by Degree Type",
       x = "Number of Records",
       y = "Degree") +
  expand_limits(x = max(degree_data$n) * 1.2)
```

```{r degree_table}
#| echo: false

degree_data |>
  arrange(desc(n)) |>
  rename(Degree = degree, Count = n, `Percentage (%)` = pct) |>
  knitr::kable(caption = "Supervision by Degree Type")
```

## Supervision Status

```{r status_distribution}
#| fig-width: 10
#| fig-height: 6

# Prepare data
status_data <- people |>
  filter(!is.na(status)) |>
  count(status, sort = TRUE) |>
  mutate(
    status = fct_reorder(status, n),
    pct = round(n / sum(n) * 100, 1)
  )

# Create plot
status_data |>
  ggplot(aes(x = n, y = status, fill = status)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(n, " (", pct, "%)")),
            hjust = -0.1, size = 4) +
  labs(title = "Distribution by Status",
       x = "Number of Records",
       y = "Status") +
  expand_limits(x = max(status_data$n) * 1.2)
```

## Yearly Trends

```{r yearly_trends}
#| fig-width: 10
#| fig-height: 6

# Prepare data - BSc and MSc only
yearly_data <- people |>
  filter(!is.na(year), degree %in% c("bsc", "msc")) |>
  count(year, degree)

if (nrow(yearly_data) > 0) {
  # Create plot with two lines
  yearly_data |>
    ggplot(aes(x = year, y = n, color = degree)) +
    geom_line(linewidth = 1) +
    geom_point(size = 3) +
    geom_text(aes(label = n), vjust = -1, size = 4, show.legend = FALSE) +
    labs(title = "BSc and MSc Supervision Records by End Year",
         x = "Year",
         y = "Number of Records",
         color = "Degree") +
    scale_x_continuous(breaks = seq(min(yearly_data$year), max(yearly_data$year), 1)) +
    scale_color_manual(values = c("bsc" = "#E69F00", "msc" = "#56B4E9"),
                       labels = c("bsc" = "BSc", "msc" = "MSc")) +
    expand_limits(y = max(yearly_data$n) * 1.15) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
} else {
  cat("No year data available")
}
```

```{r yearly_by_degree}
#| fig-width: 10
#| fig-height: 6

# Prepare data
yearly_degree_data <- people |>
  filter(!is.na(year), !is.na(degree)) |>
  count(year, degree)

if (nrow(yearly_degree_data) > 0) {
  # Create plot
  yearly_degree_data |>
    ggplot(aes(x = year, y = n, fill = degree)) +
    geom_col(position = "stack") +
    labs(title = "Supervision by Year and Degree Type",
         x = "Year",
         y = "Number of Records",
         fill = "Degree") +
    scale_x_continuous(breaks = seq(min(yearly_degree_data$year), max(yearly_degree_data$year), 1)) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
} else {
  cat("No year/degree data available")
}
```

## Supervision Type

```{r type_distribution}
#| fig-width: 10
#| fig-height: 6

# Prepare data
type_data <- people |>
  filter(!is.na(type)) |>
  count(type, sort = TRUE) |>
  mutate(
    type = fct_reorder(type, n),
    pct = round(n / sum(n) * 100, 1)
  )

# Create plot
type_data |>
  ggplot(aes(x = n, y = type, fill = type)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(n, " (", pct, "%)")),
            hjust = -0.1, size = 4) +
  labs(title = "Distribution by Supervision Type",
       x = "Number of Records",
       y = "Type") +
  expand_limits(x = max(type_data$n) * 1.2)
```

## ETH Department Distribution

```{r department_distribution}
#| fig-width: 10
#| fig-height: 6

# Prepare data
dept_data <- people |>
  filter(!is.na(eth_department)) |>
  count(eth_department, sort = TRUE) |>
  mutate(
    eth_department = fct_reorder(eth_department, n),
    pct = round(n / sum(n) * 100, 1)
  )

if (nrow(dept_data) > 0) {
  # Create plot
  dept_data |>
    ggplot(aes(x = n, y = eth_department, fill = eth_department)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = paste0(n, " (", pct, "%)")),
              hjust = -0.1, size = 4) +
    labs(title = "Distribution by ETH Department",
         x = "Number of Records",
         y = "Department") +
    expand_limits(x = max(dept_data$n) * 1.2)
} else {
  cat("No ETH department data available")
}
```

## Partner Organizations

```{r partner_distribution}
#| fig-width: 10
#| fig-height: 8

# Prepare data
partner_data <- people |>
  filter(!is.na(partner_organisation)) |>
  count(partner_organisation, sort = TRUE) |>
  mutate(
    partner_organisation = fct_reorder(partner_organisation, n),
    pct = round(n / sum(n) * 100, 1)
  )

if (nrow(partner_data) > 0) {
  # Create plot
  partner_data |>
    ggplot(aes(x = n, y = partner_organisation, fill = partner_organisation)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = paste0(n, " (", pct, "%)")),
              hjust = -0.1, size = 4) +
    labs(title = "Distribution by Partner Organization",
         x = "Number of Records",
         y = "Partner Organization") +
    expand_limits(x = max(partner_data$n) * 1.2)
} else {
  cat("No partner organization data available")
}
```

## Supervision Duration

Note: Duration analysis requires both start and end dates. Since start dates are not available in the current dataset, this section shows data availability instead.

```{r duration_analysis}
#| fig-width: 10
#| fig-height: 6

# Check data availability for duration analysis
has_start_date <- sum(!is.na(people$start_date))
has_end_date <- sum(!is.na(people$end_date))

cat(paste0("Records with start date: ", has_start_date, "/", nrow(people), "\n"))
cat(paste0("Records with end date: ", has_end_date, "/", nrow(people), "\n"))

if (has_start_date > 0 && has_end_date > 0) {
  # Calculate duration for archived supervisions
  duration_data <- people |>
    filter(!is.na(start_date), !is.na(end_date), status == "archived") |>
    mutate(
      duration_months = interval(start_date, end_date) / months(1)
    )

  if (nrow(duration_data) > 0) {
    # Create histogram
    duration_data |>
      ggplot(aes(x = duration_months)) +
      geom_histogram(binwidth = 2, fill = "steelblue", color = "white") +
      labs(title = "Distribution of Supervision Duration (Archived)",
           x = "Duration (months)",
           y = "Count") +
      theme(axis.text.x = element_text(angle = 0, hjust = 0.5))
  }
} else {
  cat("Insufficient data for duration analysis (missing start dates)")
}
```

```{r duration_by_degree}
#| fig-width: 10
#| fig-height: 6
#| eval: false

# This chunk is disabled since start_date data is not available
if (exists("duration_data") && nrow(duration_data) > 0) {
  # Prepare data
  duration_by_degree <- duration_data |>
    filter(!is.na(degree)) |>
    group_by(degree) |>
    summarise(
      n = n(),
      mean_duration = mean(duration_months, na.rm = TRUE),
      median_duration = median(duration_months, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(degree = fct_reorder(degree, mean_duration))

  # Create plot
  duration_by_degree |>
    ggplot(aes(x = mean_duration, y = degree, fill = degree)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = paste0(round(mean_duration, 1), " mo")),
              hjust = -0.1, size = 4) +
    labs(title = "Average Supervision Duration by Degree Type",
         x = "Mean Duration (months)",
         y = "Degree") +
    expand_limits(x = max(duration_by_degree$mean_duration) * 1.2)
}
```

## Data Publication Metrics

```{r data_publication}
#| fig-width: 10
#| fig-height: 6

# Calculate publication rates for archived supervisions
pub_data <- people |>
  filter(status == "archived") |>
  mutate(
    has_eth_collection = !is.na(eth_collection_link) & eth_collection_link != "NA"
  )

if (nrow(pub_data) > 0) {
  # Prepare summary data
  pub_summary <- tibble(
    metric = c("ETH Research Collection"),
    count = c(sum(pub_data$has_eth_collection)),
    total = nrow(pub_data)
  ) |>
    mutate(pct = round(count / total * 100, 1))

  # Create plot
  pub_summary |>
    ggplot(aes(x = metric, y = pct, fill = metric)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = paste0(count, "/", total, "\n(", pct, "%)")),
              vjust = -0.2, size = 4) +
    labs(title = "ETH Research Collection Links (Archived Supervisions)",
         x = "",
         y = "Percentage (%)") +
    ylim(0, 100) +
    expand_limits(y = max(pub_summary$pct) * 1.3)
} else {
  cat("No archived supervision data available")
}
```

## Status by Degree Type

```{r status_by_degree}
#| fig-width: 10
#| fig-height: 6

# Prepare data
status_degree_data <- people |>
  filter(!is.na(status), !is.na(degree)) |>
  count(degree, status)

# Create plot
status_degree_data |>
  ggplot(aes(x = degree, y = n, fill = status)) +
  geom_col(position = "dodge") +
  labs(title = "Status Distribution by Degree Type",
       x = "Degree",
       y = "Count",
       fill = "Status") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Summary Statistics

```{r summary_table}
#| echo: false

# Create summary by degree
people |>
  filter(!is.na(degree)) |>
  group_by(degree) |>
  summarise(
    Total = n(),
    Archived = sum(status == "archived", na.rm = TRUE),
    `In Progress` = sum(status == "in-progress", na.rm = TRUE),
    `With Partner` = sum(!is.na(partner_organisation)),
    `Has ETH Collection` = sum(!is.na(eth_collection_link) & eth_collection_link != "NA"),
    .groups = "drop"
  ) |>
  arrange(desc(Total)) |>
  knitr::kable(caption = "Summary Statistics by Degree Type")
```

```{r year_summary}
#| echo: false

# Create summary by year (end year)
year_summary_data <- people |>
  filter(!is.na(year))

if (nrow(year_summary_data) > 0) {
  year_summary_data |>
    group_by(year) |>
    summarise(
      Total = n(),
      Archived = sum(status == "archived", na.rm = TRUE),
      `In Progress` = sum(status == "in-progress", na.rm = TRUE),
      .groups = "drop"
    ) |>
    arrange(desc(year)) |>
    knitr::kable(caption = "Summary Statistics by End Year")
} else {
  cat("No year data available for summary")
}
```
