---
title: "GHE on ETH's Research Collection"
format:
  html:
    code-fold: true
    code-summary: "Display code"
    toc: true
    toc-depth: 4
    toc-location: left
execute:
  message: false
  warning: false
---

The [Research Collection](https://library.ethz.ch/en/researching-and-publishing/publishing-and-registering/publishing-in-the-research-collection.html) is ETH Zurich’s repository for publications and research data. It's where ETH Zurich faculty, staff and students can publish the full text of their work or openly share their research data (open access). Since the group was created in 2021, not only research staff of GHE but also students have published their work (theses, software, etc.) on the platform. 

This README must be rendered with the following command in your terminal: `quarto render analysis.qmd --to gfm --output README.md`


```{r}
library(tidyverse)
library(ggthemes)
library(kableExtra)

overview  <- read_csv("raw-data/ghe-research-collection.csv")
monthly_visits  <- read_csv("raw-data/monthly_visits.csv") |> 
    mutate(date = parse_date_time(month, orders = "my"))
```

### Overview

```{r}
paste0("Total publications: ", nrow(overview))

overview |> 
    group_by(publication_type) |> 
    count() |> 
    arrange(desc(n)) |>  
    kable(format = "html")
```

### Number of publications

```{r}
overview |> 
    group_by(year) |> 
    count() |> 
    ggplot(aes(x = year, y = n, label = n)) +
    geom_col() +
    geom_label() +
    labs(x = "",
    y = "Publications\n") +
        theme_few()
```

### Publication Types (absolute numbers)

```{r}
overview |> 
    group_by(year, publication_type_group) |> 
    count() |> 
    ggplot(aes(x = year, y = n, fill = publication_type_group)) +
    geom_col(position = "stack") +
    labs(x = "",
    y = "Publications\n",
    fill = "Publication Type") +
        theme_few()
```

### Publication Types (relative numbers)

```{r}
overview |> 
    group_by(year, publication_type_group) |> 
    count() |> 
    group_by(year) |> 
    mutate(rel_freq = n/sum(n))  |> 
    ggplot(aes(x = year, y = rel_freq, fill = publication_type_group)) +
    geom_col(position = "stack") +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(x = "",
    y = "Share\n",
    fill = "Publication Type") +
        theme_few()
```

### Licenses

```{r}
overview |> 
    group_by(year, license_short_group) |> 
    count() |> 
    ggplot(aes(x = year, y = n, fill = license_short_group)) +
    geom_col(position = "stack") +
    labs(x = "",
    y = "Publications\n",
    fill = "Publication Type") +
        theme_few()
```

### DOIs

```{r}
overview |> 
    group_by(year, doi_dummy) |> 
    count() |> 
    group_by(year) |> 
    mutate(rel_freq = n/sum(n))  |> 
    ggplot(aes(x = year, y = rel_freq, fill = doi_dummy)) +
    geom_col(position = "stack") +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(x = "",
    y = "Publications\n",
    fill = "Publication Type") +
        theme_few()
```

### Visits: Top 10

```{r}
monthly_visits |> 
    group_by(title) |> 
    summarize(total_visits = sum(visits)) |> 
    arrange(desc(total_visits)) |> 
    slice_max(n = 10, order_by = total_visits) |> 
    kable(format = "html")
```

### Total Monthly Visits Over Time

```{r}
monthly_visits |>
    group_by(date) |>
    summarize(total_visits = sum(visits)) |>
    ggplot(aes(x = date, y = total_visits)) +
    geom_line(size = 1, color = "#2c7fb8") +
    geom_point(size = 2, color = "#2c7fb8") +
    scale_x_date(date_labels = "%b %y",
        date_breaks = "1 month") +
    labs(x = "",
    y = "Total Visits\n",
    title = "Total Monthly Visits to Research Collection") +
    theme_few() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Publication Type Distribution by License

```{r}
overview |>
    group_by(publication_type_group, license_short_group) |>
    count() |>
    ggplot(aes(x = publication_type_group, y = n, fill = license_short_group)) +
    geom_col(position = "fill") +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(x = "",
    y = "Share\n",
    fill = "License") +
    theme_few() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Visits by Publication Type

```{r}
monthly_visits |>
    left_join(overview |> select(name, publication_type_group), by = c("title" = "name")) |>
    filter(!is.na(publication_type_group)) |>
    group_by(publication_type_group) |>
    summarize(total_visits = sum(visits)) |>
    arrange(desc(total_visits)) |>
    ggplot(aes(x = reorder(publication_type_group, total_visits), y = total_visits, fill = publication_type_group)) +
    geom_col(show.legend = FALSE) +
    coord_flip() +
    labs(x = "",
    y = "\nTotal Visits",
    title = "Total Visits by Publication Type") +
    theme_few()
```

### Visits: Individual items

```{r}
monthly_visits |>
    filter(title == "Life cycle assessment of household biogas digesters: A preliminary study on emissions and system interdependencies") |>
    ggplot(aes(x = date, y = visits)) +
    geom_col() +
    scale_x_date(date_labels = "%b %y",
        date_breaks = "1 month") +
    labs(x = "",
    y = "Visits\n",
    title = "Life cycle assessment of household biogas digesters: A preliminary study on emissions and system interdependencies") +
        theme_few()
```

```{r}
monthly_visits |>
    filter(title == "Land use in the Seychelles – Rethinking the Sustainability of Tourism") |>
    ggplot(aes(x = date, y = visits)) +
    geom_col() +
    scale_x_date(date_labels = "%b %y",
        date_breaks = "1 month") +
    labs(x = "",
    y = "Visits\n",
    title = "Land use in the Seychelles – Rethinking the Sustainability of Tourism") +
        theme_few()
```

```{r}
monthly_visits |>
    filter(title == "Air quality monitoring from open waste and incinerator burning in Cape Maclear, Malawi") |>
    ggplot(aes(x = date, y = visits)) +
    geom_col() +
    scale_x_date(date_labels = "%b %y",
        date_breaks = "1 month") +
    labs(x = "",
    y = "Visits\n",
    title = "Air quality monitoring from open waste and incinerator burning in Cape Maclear, Malawi") +
        theme_few()
```

```{r}
monthly_visits |>
    filter(title == "Plastic separation - A review") |>
    ggplot(aes(x = date, y = visits)) +
    geom_col() +
    scale_x_date(date_labels = "%b %y",
        date_breaks = "1 month") +
    labs(x = "",
    y = "Visits\n",
    title = "Plastic separation - A review") +
        theme_few()
```