---
title: "GHE Research Collection: Publication Patterns Analysis"
author: "GHE Open Science Team"
date: "December 9, 2025"
format:
  html:
    code-tools: true
  pdf:
    documentclass: article
---

## Overview

The [Research Collection](https://library.ethz.ch/en/researching-and-publishing/publishing-and-registering/publishing-in-the-research-collection.html) is ETH Zurich’s repository for publications and research data. It's where ETH Zurich faculty, staff and students can publish the full text of their work or openly share their research data (open access). Since the group was created in 2021, not only research staff of GHE but also students have published their work (theses, software, etc.) on the platform. 

This report analyzes publication patterns in the GHE Research Collection, examining trends in publication types, open access licensing, DOI adoption, and temporal patterns across 119 publications from 2021 to 2025.

This README must be rendered with the following command in your terminal: `quarto render analysis.qmd --to gfm --output README.md`

## Data Loading

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(kableExtra)
library(ggthemes)

# Read in the GHE research collection data
ghe_research_collection <- read_csv("raw-data/ghe-research-collection.csv")
monthly_visits <- read_csv("raw-data/monthly_visits.csv") |>
  mutate(date = parse_date_time(month, orders = "my"))
```

## Dataset Overview

```{r}
paste0("Total publications: ", nrow(ghe_research_collection))

ghe_research_collection |> 
    group_by(publication_type) |> 
    count() |> 
    arrange(desc(n)) |>  
    kable(format = "html")
```

### Number of publications

```{r}
ghe_research_collection |> 
    group_by(year) |> 
    count() |> 
    ggplot(aes(x = year, y = n, label = n)) +
    geom_col() +
    geom_label() +
    labs(x = "",
    y = "Publications\n") +
        theme_few()
```

## Publication Growth and Composition

```{r}
#| label: fig-publications-by-year
#| fig-cap: "Publications by year, colored by type group. Student papers represent the largest category overall."
#| fig-width: 8
#| fig-height: 5

ghe_research_collection |>
  count(year, publication_type_group) |>
  ggplot(aes(x = year, y = n, fill = publication_type_group)) +
  geom_col() +
  labs(
    title = "GHE Research Collection: Publications by Year",
    subtitle = "Showing growth from 2021 to 2025",
    x = "Year",
    y = "Number of Publications",
    fill = "Publication Type"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2") +
  theme(legend.position = "bottom")
```

### Publication Types (relative numbers)

```{r}
ghe_research_collection |> 
    group_by(year, publication_type_group) |> 
    count() |> 
    group_by(year) |> 
    mutate(rel_freq = n/sum(n))  |> 
    ggplot(aes(x = year, y = rel_freq, fill = publication_type_group)) +
    geom_col(position = "stack") +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(x = "",
    y = "Share\n",
    fill = "Publication Type") +
    scale_fill_brewer(palette = "Set2") +
    theme_few()
```


### Publication Type Distribution

```{r}
#| label: publication-types-table

ghe_research_collection |>
  count(publication_type_group, sort = TRUE) |>
  rename(
    `Publication Type Group` = publication_type_group,
    `Count` = n
  ) |>
  mutate(Percentage = round(100 * Count / sum(Count), 1)) |>
  knitr::kable()
```

Student papers (including Bachelor and Master theses) represent **`r round(100 * sum(ghe_research_collection$publication_type_group == "Student Paper") / nrow(ghe_research_collection), 1)`%** of the collection, followed by scientific articles at **`r round(100 * sum(ghe_research_collection$publication_type_group == "Scientific Article") / nrow(ghe_research_collection), 1)`%**.

## Open Access Licensing Trends

```{r}
#| label: fig-license-trends
#| fig-cap: "Proportion of publications by license type over time. Green indicates CC BY 4.0 (most permissive), red indicates Copyright, and gray indicates No license."
#| fig-width: 8
#| fig-height: 5

ghe_research_collection |>
  count(year, license_short_group) |>
  ggplot(aes(x = year, y = n, fill = license_short_group)) +
  geom_col(position = "fill") +
  labs(
    title = "License Adoption Trends in GHE Research Collection",
    subtitle = "Proportion of publications by license type (2021-2025)",
    x = "Year",
    y = "Proportion",
    fill = "License"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  scale_fill_manual(
    values = c(
      "CC BY 4.0" = "#2ecc71",
      "Copyright" = "#e74c3c",
      "No license" = "#95a5a6",
      "CC BY-NC 4.0" = "#f39c12",
      "CC BY-NC-ND 4.0" = "#e67e22",
      "CC BY-NC-SA 4.0" = "#d35400"
    )
  ) +
  theme(legend.position = "bottom")
```

### License Distribution Over Time

```{r}
#| label: license-by-year-table

ghe_research_collection |>
  count(year, license_short_group) |>
  pivot_wider(names_from = license_short_group, values_from = n, values_fill = 0) |>
  mutate(
    Total = rowSums(across(-year)),
    `% CC BY 4.0` = round(100 * `CC BY 4.0` / Total, 1)
  ) |>
  knitr::kable()
```

CC BY 4.0 adoption has increased from **`r ghe_research_collection |> filter(year == 2021) |> summarize(pct = round(100 * sum(license_short_group == "CC BY 4.0") / n(), 1)) |> pull(pct)`%** in 2021 to **`r ghe_research_collection |> filter(year == 2025) |> summarize(pct = round(100 * sum(license_short_group == "CC BY 4.0") / n(), 1)) |> pull(pct)`%** in 2025.

## DOI Assignment Patterns

```{r}
#| label: fig-doi-adoption
#| fig-cap: "DOI assignment by publication type group. Green indicates publications with DOIs, red indicates missing DOIs."
#| fig-width: 8
#| fig-height: 5

ghe_research_collection |>
  count(publication_type_group, doi_dummy) |>
  ggplot(aes(x = reorder(publication_type_group, n, sum), y = n, fill = doi_dummy)) +
  geom_col(position = "stack") +
  geom_text(
    aes(label = n),
    position = position_stack(vjust = 0.5),
    color = "white",
    fontface = "bold",
    size = 4
  ) +
  coord_flip() +
  labs(
    title = "DOI Assignment by Publication Type",
    subtitle = "83% of publications (99/119) have DOIs",
    x = NULL,
    y = "Number of Publications",
    fill = "DOI Status"
  ) +
  theme_minimal() +
  scale_fill_manual(values = c("Has DOI" = "#27ae60", "No DOI" = "#e74c3c")) +
  theme(legend.position = "bottom")
```

### DOI Adoption by Year

```{r}
ghe_research_collection |> 
    group_by(year, doi_dummy) |> 
    count() |> 
    group_by(year) |> 
    mutate(rel_freq = n/sum(n))  |> 
    ggplot(aes(x = year, y = rel_freq, fill = doi_dummy)) +
    geom_col(position = "stack") +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(x = "",
    y = "Publications\n",
    fill = "Publication Type") +
    scale_fill_manual(values = c("Has DOI" = "#27ae60", "No DOI" = "#e74c3c")) +    
    theme_few()
```

```{r}
#| label: doi-by-year-table

ghe_research_collection |>
  count(year, doi_dummy) |>
  pivot_wider(names_from = doi_dummy, values_from = n, values_fill = 0) |>
  mutate(
    Total = `Has DOI` + `No DOI`,
    `% with DOI` = round(100 * `Has DOI` / Total, 1)
  ) |>
  knitr::kable()
```

**Key observations:**

- Overall DOI adoption: **`r round(100 * sum(ghe_research_collection$doi_dummy == "Has DOI") / nrow(ghe_research_collection), 1)`%**
- Strong improvement: from **`r ghe_research_collection |> filter(year == 2021) |> summarize(pct = round(100 * sum(doi_dummy == "Has DOI") / n(), 1)) |> pull(pct)`%** in 2021 to **`r ghe_research_collection |> filter(year == 2025) |> summarize(pct = round(100 * sum(doi_dummy == "Has DOI") / n(), 1)) |> pull(pct)`%** in 2025
- **All student papers have DOIs** (`r sum(ghe_research_collection$publication_type_group == "Student Paper" & ghe_research_collection$doi_dummy == "Has DOI")`/`r sum(ghe_research_collection$publication_type_group == "Student Paper")`), demonstrating excellent metadata practices
- **Datasets have the lowest DOI adoption**: all `r sum(ghe_research_collection$publication_type_group == "Dataset")` datasets currently lack DOIs

## Identifying Metadata Gaps

### Publications Without Licenses

```{r}
#| label: missing-license-analysis

no_license <- ghe_research_collection |>
  filter(license_short == "No license")

cat("Publications without license:", nrow(no_license), "out of", nrow(ghe_research_collection), "\n")
```

### Publications Without DOIs

```{r}
#| label: missing-doi-analysis

no_doi <- ghe_research_collection |>
  filter(doi_dummy == "No DOI")

cat("Publications without DOI:", nrow(no_doi), "out of", nrow(ghe_research_collection), "\n")
```

### The Complete Overlap

```{r}
#| label: metadata-overlap-check

overlap_count <- sum(ghe_research_collection$license_short == "No license" & 
                     ghe_research_collection$doi_dummy == "No DOI")

cat("Publications missing BOTH license AND DOI:", overlap_count, "\n")
cat("This represents 100% of publications without licenses\n")
cat("and 100% of publications without DOIs.\n")
```

::: {.callout-warning}
## Metadata Gap Alert

**`r sum(ghe_research_collection$license_short == "No license" & ghe_research_collection$doi_dummy == "No DOI")` publications (16.8%)** lack both license information and DOIs. This represents a systematic metadata gap that affects the same set of publications.
:::

### Visualizing the Metadata Gap

```{r}
#| label: fig-missing-metadata
#| fig-cap: "Publications missing both license and DOI by year and type. These 20 publications represent a systematic metadata gap in the collection."
#| fig-width: 8
#| fig-height: 5

ghe_research_collection |>
  filter(license_short == "No license" & doi_dummy == "No DOI") |>
  count(publication_type, year) |>
  ggplot(aes(x = year, y = n, fill = publication_type)) +
  geom_col() +
  labs(
    title = "Publications Missing Both License and DOI",
    subtitle = "20 publications (16.8%) lack both metadata elements",
    x = "Year",
    y = "Number of Publications",
    fill = "Publication Type"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(legend.position = "right")
```

The perfect correlation between missing licenses and missing DOIs suggests these are not random omissions but rather a systematic gap in the metadata collection workflow. Student papers show exemplary metadata practices (100% DOI coverage), suggesting the workflow for datasets and some journal articles may need improvement.

## Research Impact and Engagement

Beyond publication metadata, analyzing how research is discovered and accessed provides insights into impact. This section examines visitor patterns across the collection.

### Top 10 Most Visited Publications

```{r}
#| label: top-10-visits

monthly_visits |>
  group_by(title) |>
  summarize(total_visits = sum(visits)) |>
  arrange(desc(total_visits)) |>
  slice_max(n = 10, order_by = total_visits) |>
  knitr::kable()
```

### Total Monthly Visits Over Time

```{r}
#| label: fig-monthly-visits
#| fig-cap: "Monthly visits to all research publications in the collection over time."
#| fig-width: 8
#| fig-height: 5

monthly_visits |>
  group_by(date) |>
  summarize(total_visits = sum(visits)) |>
  ggplot(aes(x = date, y = total_visits)) +
  geom_line(linewidth = 1, color = "#2ecc71") +
  geom_point(size = 2, color = "#2ecc71") +
  scale_x_date(date_labels = "%b %y", date_breaks = "1 month") +
  labs(
    x = "",
    y = "Total Visits",
    title = "Total Monthly Visits to Research Collection"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Visits by Publication Type

```{r}
#| label: fig-visits-by-type
#| fig-cap: "Total visits aggregated by publication type group."
#| fig-width: 8
#| fig-height: 5

monthly_visits |>
  left_join(ghe_research_collection |> select(name, publication_type_group), 
            by = c("title" = "name")) |>
  filter(!is.na(publication_type_group)) |>
  group_by(publication_type_group) |>
  summarize(total_visits = sum(visits)) |>
  arrange(desc(total_visits)) |>
  ggplot(aes(x = reorder(publication_type_group, total_visits), y = total_visits, 
             fill = publication_type_group)) +
  geom_col(show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  coord_flip() +
  labs(
    x = "",
    y = "Total Visits",
    title = "Total Visits by Publication Type"
  ) +
  theme_minimal()
```

## Key Findings

### 1. Publication Growth & Composition
- The collection has grown from **`r sum(ghe_research_collection$year == 2021)`** publications in 2021 to **`r sum(ghe_research_collection$year == 2024)`** in 2024
- Student work dominates (**`r sum(ghe_research_collection$publication_type_group == "Student Paper")` out of `r nrow(ghe_research_collection)` = `r round(100 * sum(ghe_research_collection$publication_type_group == "Student Paper") / nrow(ghe_research_collection), 1)`%**), including Master theses, Bachelor theses, and student papers
- Journal articles are the single largest type (**`r sum(ghe_research_collection$publication_type == "Journal Article")` publications, `r round(100 * sum(ghe_research_collection$publication_type == "Journal Article") / nrow(ghe_research_collection), 1)`%**)

### 2. Open Access Adoption is Strong
- **~70-86% of publications use CC BY 4.0** (the most permissive Creative Commons license) across all years
- Positive trend: 2025 shows **`r ghe_research_collection |> filter(year == 2025) |> summarize(pct = round(100 * sum(license_short_group == "CC BY 4.0") / n(), 1)) |> pull(pct)`%** CC BY 4.0 adoption vs. `r ghe_research_collection |> filter(year == 2021) |> summarize(pct = round(100 * sum(license_short_group == "CC BY 4.0") / n(), 1)) |> pull(pct)`% in 2021
- "Copyright" and "No license" publications are decreasing proportionally over time

### 3. DOI Assignment Patterns
- Overall **`r round(100 * sum(ghe_research_collection$doi_dummy == "Has DOI") / nrow(ghe_research_collection), 1)`%** have DOIs (`r sum(ghe_research_collection$doi_dummy == "Has DOI")` out of `r nrow(ghe_research_collection)`)
- Strong improvement over time: from `r ghe_research_collection |> filter(year == 2021) |> summarize(pct = round(100 * sum(doi_dummy == "Has DOI") / n(), 1)) |> pull(pct)`% in 2021 to **`r ghe_research_collection |> filter(year == 2025) |> summarize(pct = round(100 * sum(doi_dummy == "Has DOI") / n(), 1)) |> pull(pct)`%** in 2025
- All student papers have DOIs, showing excellent metadata practices
- **Datasets represent a gap**: none of the `r sum(ghe_research_collection$publication_type_group == "Dataset")` datasets have DOIs

### 4. Systematic Metadata Gap Identified
- **20 publications (16.8%) lack both license AND DOI** - a 100% overlap
- These are primarily datasets (11/20, 55%) and journal articles (6/20, 30%)
- **Zero student papers** have this metadata gap, indicating different workflows
- This suggests a systematic issue in metadata collection for certain publication types

### 5. Diverse Publication Types
- **`r n_distinct(ghe_research_collection$publication_type)`** different publication types, but dominated by 4 main categories
- Good mix of research outputs: articles, theses, datasets, reports, and conference materials

### 6. Research Impact and Engagement
- Monthly visits show consistent engagement with the collection
- Top publications receive substantial visitor traffic, demonstrating research impact
- Student papers and scientific articles drive the majority of visits
- Open access publications (especially CC BY 4.0) benefit from greater visibility

## Conclusion

The GHE Research Collection demonstrates strong commitment to open science principles, with high rates of open access licensing (CC BY 4.0) and DOI adoption that have improved significantly from 2021 to 2025. Student work forms the backbone of the collection, and the excellent metadata practices applied to these publications serve as a model for the entire collection.

However, a systematic metadata gap has been identified: 20 publications (16.8%) lack both licenses and DOIs, primarily affecting datasets and some journal articles. This 100% overlap suggests a workflow issue rather than random omissions. Addressing this gap—particularly for datasets—would significantly enhance the collection's discoverability and compliance with FAIR data principles.

The collection's visitor patterns show that research published with proper open access licensing and metadata attracts sustained engagement, underscoring the value of comprehensive metadata practices for research impact.
