---
title: "GHE Team Computational Competencies"
format:
  html:
    code-fold: true
    code-summary: "Display code"
    toc: true
    toc-depth: 4
    toc-location: left
execute:
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

This analysis explores the computational competencies and tool usage patterns of the Global Health Engineering team at ETH Zürich. The data comes from a team survey conducted in June 2025 to assess programming skills, version control knowledge, IDE usage, and adoption of modern tools like Large Language Models.

## Data Processing Pipeline

::: {style="width: 100%; max-width: 1200px;"}
```{mermaid}
%%| fig-width: 10
flowchart TD
    raw["Google Sheets<br/>(Pre-course Survey)"]

    subgraph script["data_processing.R"]
        p1["Read from Google Sheets"]
        p2["Create anonymous IDs"]
        p3["Remove non-team members"]
        p4["Select relevant columns"]
        p5["Clean column names"]
    end

    subgraph clean["clean-data/"]
        c1["computational.csv"]
    end

    raw --> p1
    p1 --> p2
    p2 --> p3
    p3 --> p4
    p4 --> p5
    p5 --> c1

    c1 --> analysis["analysis.qmd"]

    style raw fill:#e1f5ff
    style script fill:#fff4e1
    style clean fill:#e8f5e9
```
:::

The data is extracted from a Google Sheets survey, with respondents assigned anonymous IDs. Non-team members are filtered out, and relevant competency-related columns are selected and cleaned.

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(ggthemes)
library(scales)
library(knitr)

# Set theme for all plots
theme_set(theme_few())

# Load data
comp <- read_csv(here::here("competencies/clean-data/computational.csv"))
```

```{r}
#| label: data-processing
#| include: false

# Define helper function for converting experience to numeric scores
exp_score <- function(x) {
  case_when(
    str_detect(as.character(x), "never|none") ~ 0,
    str_detect(as.character(x), "few lines") ~ 1,
    str_detect(as.character(x), "couple of pages") ~ 2,
    str_detect(as.character(x), "larger pieces") ~ 3,
    TRUE ~ NA_real_
  )
}

# Define experience level ordering for programming questions
exp_levels <- c(
  "I have never programmed.",
  "I have written a few lines now and again.",
  "I have written programs for my own use that are a couple of pages long.",
  "I have written and maintained larger pieces of software."
)

# Define Git experience levels
git_levels <- c(
  "I have never used Git.",
  "I have used Git occasionally for basic tasks (e.g., cloning a repository, making commits).",
  "I have used Git regularly for my own projects, including branching and resolving conflicts.",
  "I have used Git extensively for collaborative development, including advanced workflows (e.g., pull requests, rebasing, resolving complex merge conflicts)."
)

# Define GitHub experience levels
github_levels <- c(
  "I have never used GitHub.",
  "I have used GitHub occasionally to view or download code repositories.",
  "I have used GitHub regularly to manage my own repositories (e.g., pushing code, managing issues, using README files).",
  "I have used GitHub extensively for collaborative development, including advanced features (e.g., pull requests, code reviews, managing projects, GitHub Actions)."
)

# Define CLI usage levels
cli_levels <- c(
  "I have never used the command-line interface (Terminal on Mac, Windows Terminal on Windows).",
  "I have used the CLI occasionally for basic tasks (e.g., navigating folders, running simple commands).",
  "I use the CLI regularly for my own projects, including scripting and automating tasks.",
  "I use the CLI extensively for advanced workflows, system administration, or integrating with other tools (e.g., SSH, environment management, custom scripts)."
)

# Define programming confidence levels
confidence_levels <- c(
  "I wouldn't know where to start.",
  "I could struggle through by trial and error with a lot of web searches.",
  "I could do it quickly with external help.",
  "I could do it quickly with no use of external help."
)

# Define LLM frequency levels
llm_freq_levels <- c(
  "I do not use LLMs.",
  "I use LLMs occasionally for specific tasks when needed (e.g once a week).",
  "I regularly use LLMs for various tasks to improve efficiency (e.g. once a day).",
  "I rely on LLMs for completing a wide range of tasks (e.g. several times a day)."
)

# Apply factors to all relevant columns
comp_processed <- comp |>
  mutate(
    experience_programming_general = factor(experience_programming_general, levels = exp_levels),
    experience_programming_r = factor(experience_programming_r, levels = exp_levels),
    experience_programming_python = factor(experience_programming_python, levels = exp_levels),
    experience_git = factor(experience_git, levels = git_levels),
    experience_github = factor(experience_github, levels = github_levels),
    cli_usage = factor(cli_usage, levels = cli_levels),
    programming_confidence = factor(programming_confidence, levels = confidence_levels),
    llm_usage = factor(llm_usage, levels = llm_freq_levels)
  )
```

## Data Overview

The survey was completed in June 2025 with a total of `r nrow(comp)` respondents. 

## Programming Experience

```{r}
#| label: fig-programming-combined
#| fig-cap: "Programming experience levels across General, R, and Python"
#| fig-height: 7
#| fig-width: 10

# Create long format data for combined visualization
comp_processed |>
  select(id, experience_programming_general, experience_programming_r, experience_programming_python) |>
  pivot_longer(
    cols = starts_with("experience_programming"),
    names_to = "category",
    values_to = "level"
  ) |>
  mutate(
    category = case_when(
      category == "experience_programming_general" ~ "General Programming",
      category == "experience_programming_r" ~ "R",
      category == "experience_programming_python" ~ "Python"
    ),
    category = factor(category, levels = c("General Programming", "R", "Python"))
  ) |>
  # Use complete to fill in missing combinations with 0
  count(category, level) |>
  complete(category, level, fill = list(n = 0)) |>
  # Create shortened labels for better display
  mutate(
    level_short = case_when(
      level == "I have never programmed." ~ "Never",
      level == "I have written a few lines now and again." ~ "Few lines",
      level == "I have written programs for my own use that are a couple of pages long." ~ "Couple pages",
      level == "I have written and maintained larger pieces of software." ~ "Larger software"
    ),
    level_short = factor(level_short, levels = c("Never", "Few lines", "Couple pages", "Larger software"))
  ) |>
  drop_na(level_short) |>
  ggplot(aes(x = level_short, y = n, fill = category)) +
  geom_col(position = "dodge", alpha = 0.8) +
  geom_text(aes(label = n), position = position_dodge(width = 0.9), hjust = -0.2, size = 3.5) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of respondents",
    fill = "Category",
    title = "Programming Experience Across Categories"
  ) +
  theme_few() +
  scale_fill_colorblind() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```



### Other Programming Languages

```{r}
#| label: tbl-other-languages
#| tbl-cap: "Other programming languages used by team"

# Parse multi-value field
comp |>
  select(other_languages) |> 
  mutate(langs = str_split(other_languages, ",\\s*")) |>
  unnest(langs) |> 
  count(langs, sort = TRUE) |>
  rename(Language = langs, Count = n) |>
  kable()
```

## Version Control Skills

### Git Experience

```{r}
#| label: fig-git-experience
#| fig-cap: "Git experience distribution"
#| fig-height: 5

comp_processed |>
  count(experience_git) |>
  complete(experience_git, fill = list(n = 0)) |>
  mutate(
    git_short = case_when(
      experience_git == "I have never used Git." ~ "Never",
      str_detect(experience_git, "occasionally") ~ "Occasionally",
      str_detect(experience_git, "regularly") ~ "Regularly",
      str_detect(experience_git, "extensively") ~ "Extensively"
    ),
    git_short = factor(git_short, levels = c("Never", "Occasionally", "Regularly", "Extensively"))
  ) |>
  drop_na(git_short) |>
  ggplot(aes(x = git_short, y = n)) +
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.2, size = 4) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Git Experience Levels"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### GitHub Experience

```{r}
#| label: fig-github-experience
#| fig-cap: "GitHub experience distribution"
#| fig-height: 5

comp_processed |>
  count(experience_github) |>
  complete(experience_github, fill = list(n = 0)) |>
  mutate(
    github_short = case_when(
      experience_github == "I have never used GitHub." ~ "Never",
      str_detect(experience_github, "occasionally") ~ "Occasionally",
      str_detect(experience_github, "regularly") ~ "Regularly",
      str_detect(experience_github, "extensively") ~ "Extensively"
    ),
    github_short = factor(github_short, levels = c("Never", "Occasionally", "Regularly", "Extensively"))
  ) |>
  drop_na(github_short) |>
  ggplot(aes(x = github_short, y = n)) +
  geom_col(fill = "#A23B72", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.2, size = 4) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "GitHub Experience Levels"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

## Tool Adoption

### IDE Usage

```{r}
#| label: tbl-ides-used
#| tbl-cap: "IDEs used by team members"

comp |>
  filter(!is.na(ides_used)) |>
  filter(ides_used != "None of the above") |>
  mutate(ide_list = str_split(ides_used, ",\\s*")) |>
  unnest(ide_list) |>
  count(ide_list, sort = TRUE) |>
  rename(IDE = ide_list, Count = n) |>
  kable()
```

### Command-Line Interface (CLI) Usage

```{r}
#| label: fig-cli-usage
#| fig-cap: "CLI usage patterns"
#| fig-height: 5

comp_processed |>
  count(cli_usage) |>
  complete(cli_usage, fill = list(n = 0)) |>
  mutate(
    cli_short = case_when(
      str_detect(cli_usage, "never") ~ "Never",
      str_detect(cli_usage, "occasionally") ~ "Occasionally",
      str_detect(cli_usage, "regularly") ~ "Regularly",
      str_detect(cli_usage, "extensively") ~ "Extensively"
    ),
    cli_short = factor(cli_short, levels = c("Never", "Occasionally", "Regularly", "Extensively"))
  ) |>
  drop_na(cli_short) |>
  ggplot(aes(x = cli_short, y = n)) +
  geom_col(fill = "#F18F01", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.2, size = 4) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Command-Line Interface Usage"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

## Large Language Model (LLM) Adoption

### LLM Usage Frequency

```{r}
#| label: fig-llm-usage
#| fig-cap: "Frequency of LLM usage"
#| fig-height: 5

comp_processed |>
  count(llm_usage) |>
  complete(llm_usage, fill = list(n = 0)) |>
  mutate(
    llm_short = case_when(
      str_detect(llm_usage, "do not") ~ "Do not use",
      str_detect(llm_usage, "occasionally") ~ "Occasionally",
      str_detect(llm_usage, "regularly") ~ "Regularly",
      str_detect(llm_usage, "rely") ~ "Rely heavily"
    ),
    llm_short = factor(llm_short, levels = c("Do not use", "Occasionally", "Regularly", "Rely heavily"))
  ) |>
  drop_na(llm_short) |>
  ggplot(aes(x = llm_short, y = n)) +
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.2, size = 4) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "LLM Usage Frequency"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### LLM Tools Used

```{r}
#| label: tbl-llm-tools
#| tbl-cap: "LLM tools and platforms used by team"

# Calculate total respondents for percentage calculation
total_respondents <- nrow(comp)

comp |>
  select(llm_tools_used) |>
  filter(!is.na(llm_tools_used)) |>
  # Split on comma followed by space, but not when inside parentheses
  mutate(tool_list = str_split(llm_tools_used, ",\\s*(?![^()]*\\))")) |>
  unnest(tool_list) |>
  mutate(tool_list = str_trim(tool_list)) |>  # Clean up any extra whitespace
  count(tool_list, sort = TRUE) |>
  mutate(
    percentage = round(n / total_respondents * 100, 1),
    percentage_formatted = str_glue("{percentage}%")
  ) |>
  rename(Tool = tool_list, Count = n, `% of Team` = percentage_formatted) |>
  select(Tool, Count, `% of Team`) |>
  kable()
```

## Skill Combinations & Patterns

### Skill Correlation Heatmap


### Individual Skill Profiles

```{r}
#| label: fig-skill-profiles
#| fig-cap: "Individual skill profiles showing experience levels across dimensions"
#| fig-height: 10
#| fig-width: 11

skill_matrix_with_id <- comp_processed |>
  mutate(
    prog_general = exp_score(experience_programming_general),
    prog_r = exp_score(experience_programming_r),
    prog_python = exp_score(experience_programming_python),
    git = case_when(
      str_detect(as.character(experience_git), "never") ~ 0,
      str_detect(as.character(experience_git), "occasionally") ~ 1,
      str_detect(as.character(experience_git), "regularly") ~ 2,
      str_detect(as.character(experience_git), "extensively") ~ 3,
      TRUE ~ NA_real_
    ),
    github = case_when(
      str_detect(as.character(experience_github), "never") ~ 0,
      str_detect(as.character(experience_github), "occasionally") ~ 1,
      str_detect(as.character(experience_github), "regularly") ~ 2,
      str_detect(as.character(experience_github), "extensively") ~ 3,
      TRUE ~ NA_real_
    ),
    cli = case_when(
      str_detect(as.character(cli_usage), "never") ~ 0,
      str_detect(as.character(cli_usage), "occasionally") ~ 1,
      str_detect(as.character(cli_usage), "regularly") ~ 2,
      str_detect(as.character(cli_usage), "extensively") ~ 3,
      TRUE ~ NA_real_
    ),
    respondent = as.character(str_glue("ID-{id}"))
  ) |>
  select(respondent, prog_general, prog_r, prog_python, git, github, cli)

# Create a heatmap-style visualization
skill_matrix_with_id |>
  pivot_longer(-respondent, names_to = "skill", values_to = "level") |>
  mutate(
    skill = factor(skill,
                   levels = c("prog_general", "prog_r", "prog_python", "git", "github", "cli"),
                   labels = c("Programming\nGeneral", "R", "Python", "Git", "GitHub", "CLI")),
    level_factor = factor(level,
                         levels = 0:3,
                         labels = c("None", "Beginner", "Intermediate", "Advanced"))
  ) |>
  mutate(respondent = factor(respondent, levels = unique(skill_matrix_with_id$respondent))) |>
  ggplot(aes(x = skill, y = respondent, fill = level_factor)) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_manual(
    values = c("None" = "#FFFFFF",
               "Beginner" = ggthemes::colorblind_pal()(8)[2],
               "Intermediate" = ggthemes::colorblind_pal()(8)[3],
               "Advanced" = ggthemes::colorblind_pal()(8)[4]),
    na.value = "grey90"
  ) +
  labs(
    x = "",
    y = "Respondent",
    fill = "Proficiency\nLevel",
    title = "Individual Skill Profiles Across Dimensions"
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.text.y = element_text(size = 8),
    panel.grid = element_blank()
  )
```

## Data Storage & Workflow

### Data Storage Formats

```{r}
#| label: fig-data-storage
#| fig-cap: "Preferred data storage formats"
#| fig-height: 5

comp |>
  count(data_storage_format, sort = TRUE) |>
  drop_na() |>
  mutate(data_storage_format = str_wrap(data_storage_format, 40)) |>
  ggplot(aes(x = reorder(data_storage_format, n), y = n)) +
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.3, size = 4) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Data Storage Format Preferences"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Document Writing Approach

```{r}
#| label: fig-document-approach
#| fig-cap: "Document writing approaches"
#| fig-height: 6

comp |>
  count(document_writing_approach, sort = TRUE) |>
  drop_na() |>
  mutate(document_writing_approach = str_wrap(document_writing_approach, 35)) |>
  ggplot(aes(x = reorder(document_writing_approach, n), y = n)) +
  geom_col(fill = "#A23B72", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.3, size = 4) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Document Writing Approaches"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

