---
title: "GHE Team Computational Competencies"
format:
  html:
    code-fold: true
    code-summary: "Display code"
    toc: true
    toc-depth: 4
    toc-location: left
execute:
  message: false
  warning: false
---

This analysis explores the computational competencies and tool usage patterns of the Global Health Engineering team at ETH Zürich. The data comes from a team survey conducted in June 2025 to assess programming skills, version control knowledge, IDE usage, and adoption of modern tools like Large Language Models.

## Data Processing Pipeline

::: {style="width: 100%; max-width: 1200px;"}
```{mermaid}
%%| fig-width: 10
flowchart TD
    raw["Google Sheets<br/>(Pre-course Survey)"]

    subgraph script["data_processing.R"]
        p1["Read from Google Sheets"]
        p2["Create anonymous IDs"]
        p3["Remove non-team members"]
        p4["Select relevant columns"]
        p5["Clean column names"]
    end

    subgraph clean["clean-data/"]
        c1["computational.csv"]
    end

    raw --> p1
    p1 --> p2
    p2 --> p3
    p3 --> p4
    p4 --> p5
    p5 --> c1

    c1 --> analysis["analysis.qmd"]

    style raw fill:#e1f5ff
    style script fill:#fff4e1
    style clean fill:#e8f5e9
```
:::

The data is extracted from a Google Sheets survey, with respondents assigned anonymous IDs. Non-team members are filtered out, and relevant competency-related columns are selected and cleaned.

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(ggthemes)
library(scales)
library(knitr)

# Set theme for all plots
theme_set(theme_few())

# Load data
comp <- read_csv("clean-data/computational.csv")
```

## Data Overview

The survey was completed in June 2025 with a total of `r nrow(comp)` respondents. 

## Programming Experience

### General Programming Experience

```{r}
#| label: fig-programming-general
#| fig-cap: "General programming experience levels"
#| fig-height: 5

# Define experience level ordering
exp_levels <- c(
  "I have never programmed.",
  "I have written a few lines now and again.",
  "I have written programs for my own use that are a couple of pages long.",
  "I have written and maintained larger pieces of software."
)

comp |>
  # TODO: Use fill of tidyr to make categories asnwered as 0 with 0
  mutate(experience_programming_general = factor(experience_programming_general, levels = exp_levels)) |>
  count(experience_programming_general) |>
  drop_na() |>
  ggplot(aes(x = experience_programming_general, y = n)) + # TODO: use coord_flip, so that labels are horizonal on y-axis
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = n), vjust = -0.5, size = 4) +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "General Programming Experience"
  ) +
  theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 9)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### R Programming Experience

```{r}
#| label: fig-programming-r
#| fig-cap: "R programming experience levels"
#| fig-height: 5

# TODO: see todos for first graph

comp |>
  mutate(experience_programming_r = factor(experience_programming_r, levels = exp_levels)) |>
  count(experience_programming_r) |>
  drop_na() |>
  ggplot(aes(x = experience_programming_r, y = n)) +
  geom_col(fill = "#A23B72", alpha = 0.8) +
  geom_text(aes(label = n), vjust = -0.5, size = 4) +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "R Programming Experience"
  ) +
  theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 9)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Python Programming Experience

```{r}
#| label: fig-programming-python
#| fig-cap: "Python programming experience levels"
#| fig-height: 5

comp |>
  mutate(experience_programming_python = factor(experience_programming_python, levels = exp_levels)) |>
  count(experience_programming_python) |>
  drop_na() |>
  ggplot(aes(x = experience_programming_python, y = n)) +
  geom_col(fill = "#F18F01", alpha = 0.8) +
  geom_text(aes(label = n), vjust = -0.5, size = 4) +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Python Programming Experience"
  ) +
  theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 9)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### R vs Python Experience Comparison

```{r}
#| label: fig-r-vs-python
#| fig-cap: "R vs Python experience comparison"
#| fig-height: 6
#| fig-width: 10

# Create numeric scores for experience levels
exp_score <- function(x) {
  case_when(
    str_detect(x, "never|none") ~ 0,
    str_detect(x, "few lines") ~ 1,
    str_detect(x, "couple of pages") ~ 2,
    str_detect(x, "larger pieces") ~ 3,
    TRUE ~ NA_real_
  )
}

comp |>
  mutate(
    r_score = exp_score(experience_programming_r),
    python_score = exp_score(experience_programming_python)
  ) |>
  filter(!is.na(r_score) & !is.na(python_score)) |>
  count(r_score, python_score) |>
  ggplot(aes(x = r_score, y = python_score, size = n, color = n)) +
  geom_point(alpha = 0.7) +
  scale_size_continuous(range = c(3, 15)) +
  scale_color_gradient(low = "#A23B72", high = "#F18F01") +
  scale_x_continuous(
    breaks = 0:3,
    labels = c("None", "Few lines", "Pages", "Larger")
  ) +
  scale_y_continuous(
    breaks = 0:3,
    labels = c("None", "Few lines", "Pages", "Larger")
  ) +
  labs(
    x = "\nR Experience Level",
    y = "Python Experience Level\n",
    title = "R vs Python Experience Patterns",
    size = "Count",
    color = "Count"
  ) +
  theme_few() +
  theme(legend.position = "right")
```

```{r}
#| label: tbl-r-python-crosstab
#| tbl-cap: "R vs Python experience cross-tabulation"

comp |>
  mutate(
    r_exp = case_when(
      str_detect(experience_programming_r, "never|none") ~ "None",
      str_detect(experience_programming_r, "few lines") ~ "Beginner",
      str_detect(experience_programming_r, "couple of pages") ~ "Intermediate",
      str_detect(experience_programming_r, "larger pieces") ~ "Advanced",
      TRUE ~ NA_character_
    ),
    python_exp = case_when(
      str_detect(experience_programming_python, "never|none") ~ "None",
      str_detect(experience_programming_python, "few lines") ~ "Beginner",
      str_detect(experience_programming_python, "couple of pages") ~ "Intermediate",
      str_detect(experience_programming_python, "larger pieces") ~ "Advanced",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(r_exp) & !is.na(python_exp)) |>
  count(r_exp, python_exp) |>
  pivot_wider(names_from = python_exp, values_from = n, values_fill = 0) |>
  kable()
```

### Other Programming Languages

```{r}
#| label: tbl-other-languages
#| tbl-cap: "Other programming languages used by team"

# Parse multi-value field
comp |>
  filter(!is.na(other_languages)) |>
  filter(other_languages != "None of the above") |>
  mutate(langs = str_split(other_languages, ",\\s*")) |>
  unnest(langs) |>
  count(langs, sort = TRUE) |>
  rename(Language = langs, Count = n) |>
  kable()
```

```{r}
#| label: fig-other-languages
#| fig-cap: "Most common other programming languages"
#| fig-height: 6

comp |>
  filter(!is.na(other_languages)) |>
  filter(other_languages != "None of the above") |>
  mutate(langs = str_split(other_languages, ",\\s*")) |>
  unnest(langs) |>
  count(langs, sort = TRUE) |>
  slice_max(n, n = 10) |>
  ggplot(aes(x = reorder(langs, n), y = n)) +
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Top 10 Other Programming Languages"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

## Version Control Skills

### Git Experience

```{r}
#| label: fig-git-experience
#| fig-cap: "Git experience distribution"
#| fig-height: 5

git_levels <- c(
  "I have never used Git.",
  "I have used Git occasionally for basic tasks (e.g., cloning a repository, making commits).",
  "I have used Git regularly for my own projects, including branching and resolving conflicts.",
  "I have used Git extensively for collaborative development, including advanced workflows (e.g., pull requests, rebasing, resolving complex merge conflicts)."
)

comp |>
  mutate(experience_git = factor(experience_git, levels = git_levels)) |>
  count(experience_git) |>
  drop_na() |>
  mutate(experience_git = str_wrap(experience_git, 40)) |>
  ggplot(aes(x = experience_git, y = n)) +
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = n), vjust = -0.5, size = 4) +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Git Experience Levels"
  ) +
  theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### GitHub Experience

```{r}
#| label: fig-github-experience
#| fig-cap: "GitHub experience distribution"
#| fig-height: 5

github_levels <- c(
  "I have never used GitHub.",
  "I have used GitHub occasionally to view or download code repositories.",
  "I have used GitHub regularly to manage my own repositories (e.g., pushing code, managing issues, using README files).",
  "I have used GitHub extensively for collaborative development, including advanced features (e.g., pull requests, code reviews, managing projects, GitHub Actions)."
)

comp |>
  mutate(experience_github = factor(experience_github, levels = github_levels)) |>
  count(experience_github) |>
  drop_na() |>
  mutate(experience_github = str_wrap(experience_github, 40)) |>
  ggplot(aes(x = experience_github, y = n)) +
  geom_col(fill = "#A23B72", alpha = 0.8) +
  geom_text(aes(label = n), vjust = -0.5, size = 4) +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "GitHub Experience Levels"
  ) +
  theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

## Tool Adoption

### IDE Usage

```{r}
#| label: tbl-ides-used
#| tbl-cap: "IDEs used by team members"

comp |>
  filter(!is.na(ides_used)) |>
  filter(ides_used != "None of the above") |>
  mutate(ide_list = str_split(ides_used, ",\\s*")) |>
  unnest(ide_list) |>
  count(ide_list, sort = TRUE) |>
  rename(IDE = ide_list, Count = n) |>
  kable()
```

```{r}
#| label: fig-ide-usage
#| fig-cap: "Most popular IDEs among team"
#| fig-height: 6

comp |>
  filter(!is.na(ides_used)) |>
  filter(ides_used != "None of the above") |>
  mutate(ide_list = str_split(ides_used, ",\\s*")) |>
  unnest(ide_list) |>
  count(ide_list, sort = TRUE) |>
  ggplot(aes(x = reorder(ide_list, n), y = n)) +
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of users",
    title = "IDE Popularity"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Command-Line Interface (CLI) Usage

```{r}
#| label: fig-cli-usage
#| fig-cap: "CLI usage patterns"
#| fig-height: 5

cli_levels <- c(
  "I have never used the command-line interface (Terminal on Mac, Windows Terminal on Windows).",
  "I have used the CLI occasionally for basic tasks (e.g., navigating folders, running simple commands).",
  "I use the CLI regularly for my own projects, including scripting and automating tasks.",
  "I use the CLI extensively for advanced workflows, system administration, or integrating with other tools (e.g., SSH, environment management, custom scripts)."
)

comp |>
  mutate(cli_usage = factor(cli_usage, levels = cli_levels)) |>
  count(cli_usage) |>
  drop_na() |>
  mutate(cli_usage = str_wrap(cli_usage, 35)) |>
  ggplot(aes(x = cli_usage, y = n)) +
  geom_col(fill = "#F18F01", alpha = 0.8) +
  geom_text(aes(label = n), vjust = -0.5, size = 4) +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Command-Line Interface Usage"
  ) +
  theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

## Large Language Model (LLM) Adoption

### LLM Usage Frequency

```{r}
#| label: fig-llm-usage
#| fig-cap: "Frequency of LLM usage"
#| fig-height: 5

llm_freq_levels <- c(
  "I do not use LLMs.",
  "I use LLMs occasionally for specific tasks when needed (e.g once a week).",
  "I regularly use LLMs for various tasks to improve efficiency (e.g. once a day).",
  "I rely on LLMs for completing a wide range of tasks (e.g. several times a day)."
)

comp |>
  mutate(llm_usage = factor(llm_usage, levels = llm_freq_levels)) |>
  count(llm_usage) |>
  drop_na() |>
  mutate(llm_usage = str_wrap(llm_usage, 35)) |>
  ggplot(aes(x = llm_usage, y = n)) +
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = n), vjust = -0.5, size = 4) +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "LLM Usage Frequency"
  ) +
  theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### LLM Tools Used

```{r}
#| label: tbl-llm-tools
#| tbl-cap: "LLM tools and platforms used by team"

comp |>
  filter(!is.na(llm_tools_used)) |>
  mutate(tool_list = str_split(llm_tools_used, ",\\s*")) |>
  unnest(tool_list) |>
  count(tool_list, sort = TRUE) |>
  rename(Tool = tool_list, Count = n) |>
  kable()
```

```{r}
#| label: fig-llm-tools
#| fig-cap: "Most popular LLM tools"
#| fig-height: 7

comp |>
  filter(!is.na(llm_tools_used)) |>
  mutate(tool_list = str_split(llm_tools_used, ",\\s*")) |>
  unnest(tool_list) |>
  count(tool_list, sort = TRUE) |>
  ggplot(aes(x = reorder(tool_list, n), y = n)) +
  geom_col(fill = "#A23B72", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.3, size = 3.5) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of users",
    title = "LLM Tool Adoption"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### LLM Usage by Programming Experience

```{r}
#| label: fig-llm-by-prog-exp
#| fig-cap: "LLM usage frequency by programming experience"
#| fig-height: 6
#| fig-width: 10

comp |>
  mutate(
    prog_exp = case_when(
      str_detect(experience_programming_general, "never") ~ "Never",
      str_detect(experience_programming_general, "few lines") ~ "Beginner",
      str_detect(experience_programming_general, "couple of pages") ~ "Intermediate",
      str_detect(experience_programming_general, "larger pieces") ~ "Advanced",
      TRUE ~ NA_character_
    ),
    llm_freq = case_when(
      str_detect(llm_usage, "do not use") ~ "No use",
      str_detect(llm_usage, "occasionally") ~ "Occasional",
      str_detect(llm_usage, "regularly") ~ "Regular",
      str_detect(llm_usage, "rely") ~ "Heavy",
      TRUE ~ NA_character_
    )
  ) |>
  filter(!is.na(prog_exp) & !is.na(llm_freq)) |>
  count(prog_exp, llm_freq) |>
  ggplot(aes(x = prog_exp, y = n, fill = llm_freq)) +
  geom_col(position = "dodge", alpha = 0.8) +
  labs(
    x = "\nProgramming Experience",
    y = "Number of respondents\n",
    fill = "LLM Usage",
    title = "LLM Adoption Patterns by Programming Experience"
  ) +
  theme_few() +
  scale_fill_brewer(palette = "Set2")
```

## Skill Combinations & Patterns

### Programming Confidence

```{r}
#| label: fig-programming-confidence
#| fig-cap: "Self-reported programming confidence"
#| fig-height: 5

confidence_levels <- c(
  "I wouldn't know where to start.",
  "I could struggle through by trial and error with a lot of web searches.",
  "I could do it quickly with external help.",
  "I could do it quickly with no use of external help."
)

comp |>
  mutate(programming_confidence = factor(programming_confidence, levels = confidence_levels)) |>
  count(programming_confidence) |>
  drop_na() |>
  mutate(programming_confidence = str_wrap(programming_confidence, 35)) |>
  ggplot(aes(x = programming_confidence, y = n)) +
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = n), vjust = -0.5, size = 4) +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Programming Confidence Levels"
  ) +
  theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Skill Correlation Heatmap

```{r}
#| label: fig-skill-correlation
#| fig-cap: "Correlation between different skill levels"
#| fig-height: 7
#| fig-width: 9

# Create numeric scores for all skills
skill_matrix <- comp |>
  mutate(
    prog_general = exp_score(experience_programming_general),
    prog_r = exp_score(experience_programming_r),
    prog_python = exp_score(experience_programming_python),
    git = case_when(
      str_detect(experience_git, "never") ~ 0,
      str_detect(experience_git, "occasionally") ~ 1,
      str_detect(experience_git, "regularly") ~ 2,
      str_detect(experience_git, "extensively") ~ 3,
      TRUE ~ NA_real_
    ),
    github = case_when(
      str_detect(experience_github, "never") ~ 0,
      str_detect(experience_github, "occasionally") ~ 1,
      str_detect(experience_github, "regularly") ~ 2,
      str_detect(experience_github, "extensively") ~ 3,
      TRUE ~ NA_real_
    ),
    cli = case_when(
      str_detect(cli_usage, "never") ~ 0,
      str_detect(cli_usage, "occasionally") ~ 1,
      str_detect(cli_usage, "regularly") ~ 2,
      str_detect(cli_usage, "extensively") ~ 3,
      TRUE ~ NA_real_
    ),
    llm_freq = case_when(
      str_detect(llm_usage, "do not") ~ 0,
      str_detect(llm_usage, "occasionally") ~ 1,
      str_detect(llm_usage, "regularly") ~ 2,
      str_detect(llm_usage, "rely") ~ 3,
      TRUE ~ NA_real_
    )
  ) |>
  select(prog_general, prog_r, prog_python, git, github, cli, llm_freq) |>
  drop_na()

# Compute correlation matrix
cor_matrix <- cor(skill_matrix)

# Convert to long format for ggplot
cor_long <- cor_matrix |>
  as.data.frame() |>
  rownames_to_column("var1") |>
  pivot_longer(-var1, names_to = "var2", values_to = "correlation")

# Rename variables for better labels
var_labels <- c(
  prog_general = "Programming\nGeneral",
  prog_r = "R",
  prog_python = "Python",
  git = "Git",
  github = "GitHub",
  cli = "CLI",
  llm_freq = "LLM\nUsage"
)

cor_long |>
  mutate(
    var1 = recode(var1, !!!var_labels),
    var2 = recode(var2, !!!var_labels)
  ) |>
  ggplot(aes(x = var1, y = var2, fill = correlation)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(correlation, 2)), size = 3.5) +
  scale_fill_gradient2(
    low = "#2E86AB", mid = "white", high = "#A23B72",
    midpoint = 0, limits = c(-1, 1)
  ) +
  labs(
    x = "",
    y = "",
    fill = "Correlation",
    title = "Skill Correlation Matrix"
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  ) +
  coord_fixed()
```

### Individual Skill Profiles

```{r}
#| label: fig-skill-profiles
#| fig-cap: "Individual skill profiles showing experience levels across dimensions"
#| fig-height: 10
#| fig-width: 11

skill_matrix_with_id <- comp |>
  mutate(
    prog_general = exp_score(experience_programming_general),
    prog_r = exp_score(experience_programming_r),
    prog_python = exp_score(experience_programming_python),
    git = case_when(
      str_detect(experience_git, "never") ~ 0,
      str_detect(experience_git, "occasionally") ~ 1,
      str_detect(experience_git, "regularly") ~ 2,
      str_detect(experience_git, "extensively") ~ 3,
      TRUE ~ NA_real_
    ),
    github = case_when(
      str_detect(experience_github, "never") ~ 0,
      str_detect(experience_github, "occasionally") ~ 1,
      str_detect(experience_github, "regularly") ~ 2,
      str_detect(experience_github, "extensively") ~ 3,
      TRUE ~ NA_real_
    ),
    cli = case_when(
      str_detect(cli_usage, "never") ~ 0,
      str_detect(cli_usage, "occasionally") ~ 1,
      str_detect(cli_usage, "regularly") ~ 2,
      str_detect(cli_usage, "extensively") ~ 3,
      TRUE ~ NA_real_
    ),
    respondent = str_glue("ID-{id}")
  ) |>
  select(respondent, prog_general, prog_r, prog_python, git, github, cli) |>
  drop_na()

# Create a heatmap-style visualization
skill_matrix_with_id |>
  pivot_longer(-respondent, names_to = "skill", values_to = "level") |>
  mutate(
    skill = factor(skill,
                   levels = c("prog_general", "prog_r", "prog_python", "git", "github", "cli"),
                   labels = c("Programming\nGeneral", "R", "Python", "Git", "GitHub", "CLI"))
  ) |>
  ggplot(aes(x = skill, y = respondent, fill = level)) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_gradient2(
    low = "#FFFFFF", mid = "#A23B72", high = "#2E86AB",
    midpoint = 1.5, limits = c(0, 3),
    breaks = 0:3,
    labels = c("None", "Beginner", "Intermediate", "Advanced")
  ) +
  labs(
    x = "",
    y = "Respondent",
    fill = "Proficiency\nLevel",
    title = "Individual Skill Profiles Across Dimensions"
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    axis.text.y = element_text(size = 8),
    panel.grid = element_blank()
  )
```

### Team Skill Segments

```{r}
#| label: fig-skill-segments
#| fig-cap: "Team skill segments based on overall experience"
#| fig-height: 6

comp |>
  mutate(
    segment = case_when(
      str_detect(experience_programming_general, "never|few lines") &
        str_detect(experience_git, "never|occasionally") ~ "Beginners",
      str_detect(experience_programming_general, "couple of pages") &
        str_detect(experience_git, "regularly|occasionally") ~ "Intermediate",
      str_detect(experience_programming_general, "larger pieces") &
        str_detect(experience_git, "extensively|regularly") ~ "Advanced",
      TRUE ~ "Mixed"
    )
  ) |>
  count(segment) |>
  mutate(
    percentage = round(n / sum(n) * 100, 1),
    label = str_glue("{segment}\n(n={n}, {percentage}%)")
  ) |>
  ggplot(aes(x = "", y = n, fill = segment)) +
  geom_col(width = 1, alpha = 0.8) +
  geom_text(aes(label = str_glue("{n}\n({percentage}%)")),
            position = position_stack(vjust = 0.5),
            size = 4) +
  coord_polar(theta = "y") +
  labs(
    title = "Team Skill Segments",
    fill = "Segment"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.position = "bottom"
  ) +
  scale_fill_brewer(palette = "Set2")
```

```{r}
#| label: tbl-segment-characteristics
#| tbl-cap: "Characteristics of team skill segments"

comp |>
  mutate(
    segment = case_when(
      str_detect(experience_programming_general, "never|few lines") &
        str_detect(experience_git, "never|occasionally") ~ "Beginners",
      str_detect(experience_programming_general, "couple of pages") &
        str_detect(experience_git, "regularly|occasionally") ~ "Intermediate",
      str_detect(experience_programming_general, "larger pieces") &
        str_detect(experience_git, "extensively|regularly") ~ "Advanced",
      TRUE ~ "Mixed"
    ),
    r_exp = exp_score(experience_programming_r),
    python_exp = exp_score(experience_programming_python),
    uses_llm = !str_detect(llm_usage, "do not use")
  ) |>
  group_by(segment) |>
  summarise(
    count = n(),
    avg_r_exp = round(mean(r_exp, na.rm = TRUE), 2),
    avg_python_exp = round(mean(python_exp, na.rm = TRUE), 2),
    pct_llm_users = round(sum(uses_llm, na.rm = TRUE) / n() * 100, 1)
  ) |>
  arrange(desc(count)) |>
  kable(col.names = c("Segment", "Count", "Avg R Exp", "Avg Python Exp", "% LLM Users"))
```

## Data Storage & Workflow

### Data Storage Formats

```{r}
#| label: fig-data-storage
#| fig-cap: "Preferred data storage formats"
#| fig-height: 5

comp |>
  count(data_storage_format, sort = TRUE) |>
  drop_na() |>
  mutate(data_storage_format = str_wrap(data_storage_format, 40)) |>
  ggplot(aes(x = reorder(data_storage_format, n), y = n)) +
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.3, size = 4) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Data Storage Format Preferences"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Document Writing Approach

```{r}
#| label: fig-document-approach
#| fig-cap: "Document writing approaches"
#| fig-height: 6

comp |>
  count(document_writing_approach, sort = TRUE) |>
  drop_na() |>
  mutate(document_writing_approach = str_wrap(document_writing_approach, 35)) |>
  ggplot(aes(x = reorder(document_writing_approach, n), y = n)) +
  geom_col(fill = "#A23B72", alpha = 0.8) +
  geom_text(aes(label = n), hjust = -0.3, size = 4) +
  coord_flip() +
  labs(
    x = "",
    y = "\nNumber of respondents",
    title = "Document Writing Approaches"
  ) +
  theme_few() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

## Summary

```{r}
#| label: summary-calculations
#| echo: false

total_respondents <- nrow(comp)

advanced_programmers <- comp |>
  filter(str_detect(experience_programming_general, "larger pieces")) |>
  nrow()

r_users <- comp |>
  filter(!str_detect(experience_programming_r, "none")) |>
  nrow()

python_users <- comp |>
  filter(!str_detect(experience_programming_python, "none")) |>
  nrow()

git_advanced <- comp |>
  filter(str_detect(experience_git, "extensively")) |>
  nrow()

llm_users <- comp |>
  filter(!str_detect(llm_usage, "do not use")) |>
  nrow()

llm_heavy <- comp |>
  filter(str_detect(llm_usage, "rely")) |>
  nrow()

top_ide <- comp |>
  filter(!is.na(ides_used)) |>
  mutate(ide_list = str_split(ides_used, ",\\s*")) |>
  unnest(ide_list) |>
  count(ide_list, sort = TRUE) |>
  slice_max(n, n = 1) |>
  pull(ide_list)

top_llm <- comp |>
  filter(!is.na(llm_tools_used)) |>
  mutate(tool_list = str_split(llm_tools_used, ",\\s*")) |>
  unnest(tool_list) |>
  count(tool_list, sort = TRUE) |>
  slice_max(n, n = 1) |>
  pull(tool_list)
```

This analysis examined computational competencies of **`r total_respondents`** team members at Global Health Engineering, based on a survey conducted in June 2025.

### Key Findings:

**Programming Skills:**

- **`r advanced_programmers`** team members (`r round(advanced_programmers/total_respondents*100, 1)`%) have written and maintained larger pieces of software
- **`r r_users`** members (`r round(r_users/total_respondents*100, 1)`%) have some R experience
- **`r python_users`** members (`r round(python_users/total_respondents*100, 1)`%) have some Python experience
- R experience is more common than Python experience overall

**Version Control:**

- **`r git_advanced`** members (`r round(git_advanced/total_respondents*100, 1)`%) use Git extensively for collaborative development
- Strong positive correlation between Git and GitHub experience (r = `r round(cor(skill_matrix$git, skill_matrix$github), 2)`)

**Tool Adoption:**

- Most popular IDE: **`r top_ide`**
- **`r llm_users`** members (`r round(llm_users/total_respondents*100, 1)`%) use LLMs for their work
- **`r llm_heavy`** members (`r round(llm_heavy/total_respondents*100, 1)`%) rely on LLMs several times daily
- Most popular LLM tool: **`r top_llm`**

**Interesting Patterns:**

- LLM adoption is high across all programming experience levels
- Strong correlation between general programming experience and both R/Python skills
- Team shows diversity in skill levels, enabling peer learning opportunities
- Majority prefer machine-readable data formats (CSV/JSON) over spreadsheets
