---
title: "GHE GitHub Repository Analysis"
author: "Global Health Engineering"
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
execute:
  echo: true
  warning: false
  message: false
code-fold: true
---

```{r}
#| label: setup
library(tidyverse)
library(scales)
library(knitr)
library(ggthemes)

# Set theme for all plots
theme_set(theme_few())
```

## Repositories

```{r}
github_repos <- read_csv("clean-data/github_repos.csv")
```

### Private/public repositories

```{r}
github_repos |>
     count(created_year, is_private) |>
     ggplot(aes(x = created_year, y = n, fill = is_private)) +
     geom_col(position = position_dodge2()) +
     labs(
          x = "",
          y = "Number of repositories\n",
          fill = "Visibility"
     ) +
     theme_few() +
     theme(panel.grid = element_blank())

```

### Licenses

```{r}
github_repos |>
     count(created_year, license_dummy) |>
     ggplot(aes(x = created_year, y = n, fill = license_dummy)) +
     geom_col(position = position_dodge2()) +
     labs(
          x = "",
          y = "Number of repositories\n",
          fill = "License?"
     ) +
     theme_few() +
     theme(panel.grid = element_blank())
```

### Main repo languages over the years

```{r}
github_repos |>
     filter(publication_type != "Student paper") |>
     count(created_year, primary_language_ext2) |>
     ggplot(aes(x = created_year, y = n, fill = primary_language_ext2)) +
     geom_col(position = position_dodge2()) +
     labs(
          x = "",
          y = "Number of repositories\n",
          fill = "Main repo language",
          subtitle = "Student theses excluded (no primary language available)"
     ) +
     theme_bw() +
     theme(panel.grid = element_blank())
```

```{r}
github_repos |>
     filter(publication_type != "Student paper") |>
     count(created_year, primary_language_ext2) |> 
  group_by(created_year) |> 
  mutate(rel_freq = n/sum(n)) |> 
  ggplot(aes(x = created_year, y = rel_freq, fill = primary_language_ext2)) +
     geom_col(position = "stack") +
  scale_y_continuous(labels = scales::label_percent()) +
  labs(y = "",
       x = "",
       fill = "Main repo language")
```


### Publication Types over the years

```{r}
github_repos |>
     count(created_year, publication_type) |>
     ggplot(aes(x = created_year, y = n, fill = publication_type)) +
     geom_col(position = position_dodge2()) +
     labs(x = "", y = "Number of repositories\n", fill = "Publication Type") +
     theme_bw() +
     theme(panel.grid = element_blank())
```


## Contributors

```{r}
#| label: load-data
contributors <- read_csv("raw-data/contributors.csv")
```


### Top contributors by total commits

```{r}
#| label: top-contributors-total
top_contributors <- contributors %>%
  group_by(login) %>%
  summarise(
    total_contributions = sum(contributions),
    repos_contributed_to = n_distinct(repository)
  ) %>%
  arrange(desc(total_contributions)) %>%
  head(10)

# Display table
kable(top_contributors, 
      caption = "Top 10 Contributors by Total Contributions",
      col.names = c("Contributor", "Total Contributions", "Repositories"))
```

```{r}
#| label: fig-top-contributors
#| fig-cap: "Top 10 Contributors by Total Contributions Across All Repositories"
#| fig-height: 6
#| fig-width: 10

ggplot(top_contributors, aes(x = reorder(login, total_contributions), 
                              y = total_contributions)) +
  geom_col(fill = "#2E86AB", alpha = 0.8) +
  geom_text(aes(label = total_contributions), 
            hjust = -0.2, 
            size = 3.5) +
  coord_flip() +
  labs(
    x = "Contributor",
    y = "Total Number of Contributions",
    caption = paste("Data from", n_distinct(contributors$repository), "repositories")
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                     labels = comma)
```

### Average Number of Contributors per Repository

```{r}
#| label: avg-contributors
repo_stats <- contributors %>%
  group_by(repository) %>%
  summarise(
    num_contributors = n_distinct(login),
    total_contributions = sum(contributions)
  )

avg_contributors <- mean(repo_stats$num_contributors)
median_contributors <- median(repo_stats$num_contributors)

# Create summary table
summary_stats <- tibble(
  Metric = c("Average Contributors per Repository", 
             "Median Contributors per Repository",
             "Total Unique Contributors",
             "Total Repositories"),
  Value = c(round(avg_contributors, 2),
            median_contributors,
            n_distinct(contributors$login),
            n_distinct(contributors$repository))
)

kable(summary_stats, caption = "Repository Contributor Statistics")
```

```{r}
#| label: fig-contributor-distribution
#| fig-cap: "Distribution of Number of Contributors per Repository"
#| fig-height: 5
#| fig-width: 8

ggplot(repo_stats, aes(x = num_contributors)) +
  geom_histogram(binwidth = 1, fill = "#2E86AB", alpha = 0.8, color = "white") +
  geom_vline(xintercept = avg_contributors, 
             color = "red", linetype = "dashed", size = 1) +
  geom_vline(xintercept = median_contributors, 
             color = "orange", linetype = "dashed", size = 1) +
  annotate("text", x = avg_contributors + 0.5, y = Inf, 
           label = paste("Mean =", round(avg_contributors, 1)), 
           hjust = 0, vjust = 2, color = "red", fontface = "bold") +
  annotate("text", x = median_contributors + 0.5, y = Inf, 
           label = paste("Median =", median_contributors), 
           hjust = 0, vjust = 3.5, color = "orange", fontface = "bold") +
  labs(
    title = "Distribution of Contributors per Repository",
    x = "Number of Contributors",
    y = "Number of Repositories",
    subtitle = paste("Based on", n_distinct(repo_stats$repository), "repositories")
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12)
  )
```

### Contributors by Number of Repositories

```{r}
#| label: contributors-by-repo-count
contributors_by_repos <- contributors %>%
  group_by(login) %>%
  summarise(
    num_repositories = n_distinct(repository),
    total_contributions = sum(contributions),
    avg_contributions_per_repo = round(mean(contributions), 1)
  ) %>%
  arrange(desc(num_repositories)) %>%
  head(10)

kable(contributors_by_repos,
      caption = "Top 10 Contributors by Number of Repositories",
      col.names = c("Contributor", "Repositories", "Total Contributions", "Avg per Repo"))
```

```{r}
#| label: fig-contributors-by-repos
#| fig-cap: "Top 10 Contributors by Number of Repositories"
#| fig-height: 6
#| fig-width: 10

ggplot(contributors_by_repos, aes(x = reorder(login, num_repositories), 
                                   y = num_repositories)) +
  geom_col(fill = "#A23B72", alpha = 0.8) +
  geom_text(aes(label = num_repositories), 
            hjust = -0.2, 
            size = 3.5) +
  coord_flip() +
  labs(
    title = "Top 10 Contributors by Repository Count",
    subtitle = "Contributors who have contributed to the most repositories",
    x = "Contributor",
    y = "Number of Repositories",
    caption = paste("Total repositories analyzed:", n_distinct(contributors$repository))
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))
```

### Contribution Concentration Analysis

```{r}
#| label: contribution-concentration
# Calculate cumulative contribution percentage
contribution_concentration <- contributors %>%
  group_by(login) %>%
  summarise(total_contributions = sum(contributions)) %>%
  arrange(desc(total_contributions)) %>%
  mutate(
    cumulative_contributions = cumsum(total_contributions),
    cumulative_percentage = cumulative_contributions / sum(total_contributions) * 100,
    contributor_rank = row_number()
  )

# Find how many contributors make up 50% and 80% of contributions
contributors_for_50 <- min(which(contribution_concentration$cumulative_percentage >= 50))
contributors_for_80 <- min(which(contribution_concentration$cumulative_percentage >= 80))

cat(paste("Top", contributors_for_50, "contributors (", 
          round(contributors_for_50/n_distinct(contributors$login)*100, 1), 
          "% of people) make 50% of all contributions\n"))
cat(paste("Top", contributors_for_80, "contributors (",
          round(contributors_for_80/n_distinct(contributors$login)*100, 1),
          "% of people) make 80% of all contributions\n"))
```

```{r}
#| label: fig-lorenz-curve
#| fig-cap: "Lorenz Curve of Contribution Distribution"
#| fig-height: 6
#| fig-width: 8

# Prepare data for Lorenz curve
lorenz_data <- contribution_concentration %>%
  mutate(
    contributor_percentage = contributor_rank / max(contributor_rank) * 100
  ) %>%
  add_row(contributor_percentage = 0, cumulative_percentage = 0, .before = 1)

ggplot(lorenz_data, aes(x = contributor_percentage, y = cumulative_percentage)) +
  geom_line(size = 1.5, color = "#2E86AB") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50") +
  geom_ribbon(aes(ymin = contributor_percentage, ymax = cumulative_percentage), 
              fill = "#2E86AB", alpha = 0.2) +
  annotate("text", x = 50, y = 25, label = "Perfect Equality", 
           angle = 45, color = "gray50", size = 4) +
  annotate("point", x = contributors_for_50/n_distinct(contributors$login)*100, 
           y = 50, color = "red", size = 3) +
  annotate("text", x = contributors_for_50/n_distinct(contributors$login)*100 + 5, 
           y = 50, label = paste(contributors_for_50, "people\nmake 50% of\ncontributions"),
           size = 3, hjust = 0, color = "red") +
  labs(
    title = "Contribution Inequality (Lorenz Curve)",
    subtitle = "Distribution of contributions across all contributors",
    x = "Cumulative % of Contributors",
    y = "Cumulative % of Contributions"
  ) +
  theme_minimal() +
  scale_x_continuous(breaks = seq(0, 100, 20)) +
  scale_y_continuous(breaks = seq(0, 100, 20))
```

### Time-Based Insights (Repository Age Analysis)

```{r, include=FALSE, eval=FALSE}
#| label: repo-freshness
# Load repos.csv to get creation dates
repos_info <- read_delim("repos.csv", delim = ";", show_col_types = FALSE)

# Parse dates and calculate age
repos_with_age <- repos_info %>%
  mutate(
    createdAt = as.Date(createdAt),
    updatedAt = as.Date(updatedAt),
    age_days = as.numeric(Sys.Date() - createdAt),
    days_since_update = as.numeric(Sys.Date() - updatedAt)
  ) %>%
  select(name, createdAt, updatedAt, age_days, days_since_update)

# Join with contributor data
repo_activity <- repo_stats %>%
  left_join(repos_with_age, by = c("repository" = "name")) %>%
  filter(!is.na(age_days))

# Correlation between age and contributors
cor_age_contributors <- cor(repo_activity$age_days, repo_activity$num_contributors, use = "complete.obs")

cat(paste("Correlation between repository age and number of contributors:", 
          round(cor_age_contributors, 3), "\n"))

# Recently active repositories
recent_repos <- repo_activity %>%
  filter(days_since_update < 30) %>%
  arrange(desc(num_contributors)) %>%
  head(5)

cat("\nMost collaborative repositories updated in the last 30 days:\n")
kable(recent_repos %>% select(repository, num_contributors, days_since_update),
      col.names = c("Repository", "Contributors", "Days Since Update"))
```

```{r, include=FALSE, eval=FALSE}
#| label: fig-age-vs-contributors
#| fig-cap: "Repository Age vs Number of Contributors"
#| fig-height: 6
#| fig-width: 10

ggplot(repo_activity, aes(x = age_days/365, y = num_contributors)) +
  geom_point(aes(size = total_contributions, color = days_since_update < 90), 
             alpha = 0.6) +
  geom_smooth(method = "loess", se = TRUE, color = "#2E86AB") +
  scale_size_continuous(name = "Total\nContributions", range = c(2, 10)) +
  scale_color_manual(values = c("gray50", "#F18F01"), 
                     labels = c("Inactive (90+ days)", "Active (<90 days)"),
                     name = "Activity Status") +
  labs(
    title = "Repository Maturity vs Collaboration",
    subtitle = "How repository age relates to number of contributors",
    x = "Repository Age (years)",
    y = "Number of Contributors"
  ) +
  theme_minimal()
```

## Summary

This analysis examined `r n_distinct(contributors$repository)` repositories with a total of `r n_distinct(contributors$login)` unique contributors. 

### Key findings:

- The average repository has **`r round(avg_contributors, 1)` contributors**
- The top contributor (`r top_contributors$login[1]`) has made **`r format(top_contributors$total_contributions[1], big.mark = ",")` contributions** across all repositories
- The most collaborative contributor (`r contributors_by_repos$login[1]`) has contributed to **`r contributors_by_repos$num_repositories[1]` different repositories**

### Interesting patterns discovered:

- **Contribution inequality**: Just `r contributors_for_50` people (`r round(contributors_for_50/n_distinct(contributors$login)*100, 1)`% of all contributors) account for 50% of all contributions

```{r}
#- **Repository age correlation**: The correlation between repository age and number of contributors is `r round(cor_age_contributors, 2)`, suggesting `r ifelse(cor_age_contributors > 0.3, "older repositories tend to attract more contributors", "age doesn't strongly predict contributor count")`

# - **Repository collaboration**: `r round(multi_repos/nrow(repo_stats)*100, 1)`% of repositories have multiple contributors

#- **Power contributors**: `r sum(contributor_categories$category == "Power Contributors (100+ commits)")` individuals have made 100+ contributions each
```

