---
title: "Zenodo Research Group Publications Analysis"
format: html
editor: source
execute:
  echo: false
---

```{r extract_data}
#| include: false
#| eval: false

library(here)
# Set to eval: true to extract and transform data before rendering
# This chunk extracts data from Zenodo API and then transforms it

# Extract data 
result_extract <- system2(
  "Rscript",
  args = c(here("zenodo", "scripts", "01_extract_zenodo_data.R")),
  stdout = TRUE,
  stderr = TRUE
)

if (!is.null(attr(result_extract, "status")) && attr(result_extract, "status") != 0) {
  cat("Error extracting data:\n")
  cat(paste(result_extract, collapse = "\n"), "\n")
  stop("Failed to extract Zenodo data")
}

cat(paste(result_extract, collapse = "\n"), "\n\n")

# Transform the extracted data
result_transform <- system2(
  "Rscript",
  args = c(here("zenodo", "scripts", "02_transform_zenodo_data.R")),
  stdout = TRUE,
  stderr = TRUE
)

if (!is.null(attr(result_transform, "status")) && attr(result_transform, "status") != 0) {
  cat("Error transforming data:\n")
  cat(paste(result_transform, collapse = "\n"), "\n")
  stop("Failed to transform Zenodo data")
}

cat(paste(result_transform, collapse = "\n"), "\n")
```

```{r setup}
#| include: false
library(tidyverse)
library(lubridate)
library(scales)
library(ggthemes)
library(here)

# Load data
load(here("zenodo", "data", "zenodo_ghe_data.rda"))

# Set theme
theme_set(theme_minimal())
```

## Overview

This analysis covers **`r nrow(zenodo_data)` publications** from the research group, spanning from **`r format(min(zenodo_data$publication_date), "%B %d, %Y")`** to **`r format(max(zenodo_data$publication_date), "%B %d, %Y")`**.

**Key Statistics:**

- Total records: `r nrow(zenodo_data)`
- Unique authors: `r n_distinct(authors_data$author_name)`
- Date range: `r min(zenodo_data$publication_date)` to `r max(zenodo_data$publication_date)`

## Publication Trends Over Time

```{r trends}
#| fig-width: 10
#| fig-height: 6

# Prepare data
trends_data <- zenodo_data |>
  mutate(year_month = floor_date(publication_date, "month")) |>
  count(year_month)

# Create plot
trends_data |>
  ggplot(aes(x = year_month, y = n)) +
  geom_line(color = colorblind_pal()(8)[1], linewidth = 1) +
  geom_point(color = colorblind_pal()(8)[1], size = 2) +
  labs(title = "Publications Over Time",
       x = "Date",
       y = "Number of Publications") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r trends_by_type}
#| fig-width: 10
#| fig-height: 6

# Prepare data
trends_by_type_data <- zenodo_data |>
  filter(!is.na(resource_type)) |>
  mutate(year = year(publication_date)) |>
  count(year, resource_type)

# Create plot
trends_by_type_data |>
  ggplot(aes(x = year, y = n, fill = resource_type)) +
  geom_col(position = "stack") +
  labs(title = "Publications by Type Over Time",
       x = "Year",
       y = "Number of Publications",
       fill = "Resource Type") +
  scale_fill_colorblind() +
  scale_x_continuous(breaks = \(x) seq(floor(min(x)), ceiling(max(x)), 1))
```

## Resource Types

```{r resource_types}
#| fig-width: 10
#| fig-height: 6

# Prepare data
resource_types_data <- zenodo_data |>
  filter(!is.na(resource_type)) |>
  count(resource_type, sort = TRUE) |>
  mutate(resource_type = fct_reorder(resource_type, n))

# Create plot
resource_types_data |>
  ggplot(aes(x = n, y = resource_type, fill = resource_type)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Distribution of Resource Types",
       x = "Number of Publications",
       y = "Resource Type") +
  scale_fill_colorblind()
```

## Licenses

```{r license_table}
#| echo: false

license_data <- zenodo_data |>
  filter(!is.na(license_id))

if (nrow(license_data) > 0) {
  license_data |>
    count(license_id, license_title, sort = TRUE) |>
    mutate(percentage = round(n / sum(n) * 100, 1)) |>
    rename(`License ID` = license_id,
           `License Name` = license_title,
           `Count` = n,
           `Percentage` = percentage) |>
    knitr::kable(caption = "License Distribution")
}
```

## Collaboration Patterns

```{r collaboration}
#| fig-width: 10
#| fig-height: 6

# Prepare data
collaboration_data <- zenodo_data |>
  left_join(authors_data |> count(record_id, name = "n_authors"),
            by = "record_id") |>
  filter(!is.na(n_authors)) |>
  count(n_authors)

# Create plot
collaboration_data |>
  ggplot(aes(x = n_authors, y = n)) +
  geom_col(fill = colorblind_pal()(8)[2]) +
  labs(title = "Number of Authors per Publication",
       x = "Number of Authors",
       y = "Number of Publications") +
  scale_x_continuous(breaks = \(x) seq(floor(min(x)), ceiling(max(x)), 1))
```

```{r top_authors}
#| fig-width: 10
#| fig-height: 8

# Prepare data
auth_count <- authors_data |>
  count(author_name, sort = TRUE) |>
  slice_head(n = 20) |>
  mutate(author_name = fct_reorder(author_name, n))

# Recycle colorblind palette colors for all authors
n_auth <- nrow(auth_count)
colors <- rep(colorblind_pal()(8), length.out = n_auth)

# Create plot
auth_count |>
  ggplot(aes(x = n, y = author_name, fill = author_name)) +
  geom_col(show.legend = FALSE) +
  labs(title = "Top 20 Most Prolific Authors",
       x = "Number of Publications",
       y = "Author") +
  scale_fill_manual(values = colors)
```

## Topics

```{r subjects}
#| fig-width: 10
#| fig-height: 8

if (nrow(subjects_data) > 0) {
  # Prepare data
  subj_count <- subjects_data |>
    count(subject, sort = TRUE) |>
    slice_head(n = 20) |>
    mutate(subject = fct_reorder(subject, n))

  # Recycle colorblind palette colors for all subjects
  n_subj <- nrow(subj_count)
  colors <- rep(colorblind_pal()(8), length.out = n_subj)

  # Create plot
  subj_count |>
    ggplot(aes(x = n, y = subject, fill = subject)) +
    geom_col(show.legend = FALSE) +
    labs(title = "Top 20 Most Common Subjects",
         x = "Frequency",
         y = "Subject") +
    scale_fill_manual(values = colors)
} else {
  cat("No subject data available")
}
```

## Communities

```{r community_distribution}
#| fig-width: 10
#| fig-height: 6

# Calculate communities per record
communities_per_record <- zenodo_data |>
  left_join(
    communities_data |> count(record_id, name = "n_communities"),
    by = "record_id"
  ) |>
  mutate(n_communities = replace_na(n_communities, 0))

# Prepare data for plotting
community_dist_plot_data <- communities_per_record |>
  count(n_communities) |>
  mutate(
    pct = round(n / sum(n) * 100, 1),
    community_label = case_when(
      n_communities == 0 ~ "No community",
      n_communities == 1 ~ "1 community",
      TRUE ~ paste(n_communities, "communities")
    ),
    community_label = fct_reorder(community_label, n_communities)
  )

# Create plot
community_dist_plot_data |>
  ggplot(aes(x = n, y = community_label, fill = as.factor(n_communities))) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(n, " (", pct, "%)")),
            hjust = -0.2, size = 4) +
  labs(title = "Distribution of Community Membership",
       subtitle = "Number of communities per record",
       x = "Number of Records",
       y = "Community Status") +
  scale_fill_colorblind() +
  expand_limits(x = max(community_dist_plot_data$n) * 1.15)
```

```{r community_stats_calc}
#| echo: false
#| include: false

records_in_communities <- nrow(zenodo_data |> filter(record_id %in% communities_data$record_id))
records_not_in_communities <- nrow(zenodo_data) - records_in_communities
pct_in_communities <- round(records_in_communities / nrow(zenodo_data) * 100, 1)
```

**Community Membership Statistics:**

- Records in communities: `r records_in_communities` / `r nrow(zenodo_data)` (`r pct_in_communities`%)
- Records not in communities: `r records_not_in_communities`
- Unique communities: `r n_distinct(communities_data$community_id)`

```{r community_table}
#| echo: false

if (nrow(communities_data) > 0) {
  communities_data |>
    count(community_slug, community_title, sort = TRUE) |>
    rename(Community = community_slug, Title = community_title, Records = n) |>
    knitr::kable(caption = "Communities")
}
```

## Related Identifiers

```{r related_identifiers}
#| fig-width: 10
#| fig-height: 6

if (nrow(related_identifiers_data) > 0) {
  # Prepare data
  related_identifiers_plot_data <- related_identifiers_data |>
    filter(!is.na(relation_type)) |>
    count(relation_type, sort = TRUE) |>
    mutate(relation_type = fct_reorder(relation_type, n))

  # Create plot
  related_identifiers_plot_data |>
    ggplot(aes(x = n, y = relation_type, fill = relation_type)) +
    geom_col(show.legend = FALSE) +
    labs(title = "Types of Related Identifiers",
         x = "Frequency",
         y = "Relation Type") +
    scale_fill_colorblind()
} else {
  cat("No related identifiers data available")
}
```

## Metadata Completeness

```{r metadata_stats_calc}
#| echo: false
#| include: false

orcid_stats <- authors_data |>
  summarise(
    total_authors = n(),
    with_orcid = sum(!is.na(author_orcid)),
    pct_orcid = round(with_orcid / total_authors * 100, 1),
    unique_orcids = n_distinct(author_orcid[!is.na(author_orcid)])
  )

affiliation_stats <- authors_data |>
  summarise(
    total_authors = n(),
    with_affiliation = sum(!is.na(author_affiliation)),
    pct_affiliation = round(with_affiliation / total_authors * 100, 1)
  )
```

**ORCID Coverage:**

- Authors with ORCID: `r orcid_stats$with_orcid` / `r orcid_stats$total_authors` (`r orcid_stats$pct_orcid`%)
- Unique ORCIDs: `r orcid_stats$unique_orcids`

**Affiliation Coverage:**

- Authors with affiliation: `r affiliation_stats$with_affiliation` / `r affiliation_stats$total_authors` (`r affiliation_stats$pct_affiliation`%)

```{r metadata_completeness_plot}
#| fig-width: 8
#| fig-height: 5

# Prepare data
metadata_completeness_data <- tibble(
  metric = c("ORCID", "Affiliation"),
  percentage = c(
    sum(!is.na(authors_data$author_orcid)) / nrow(authors_data) * 100,
    sum(!is.na(authors_data$author_affiliation)) / nrow(authors_data) * 100
  )
)

# Create plot
metadata_completeness_data |>
  ggplot(aes(x = metric, y = percentage, fill = metric)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")),
            vjust = -0.5, size = 5) +
  labs(title = "Author Metadata Completeness",
       x = "Metadata Field",
       y = "Percentage of Authors") +
  scale_fill_colorblind() +
  ylim(0, 100)
```

## Affiliation 

```{r top_affiliations}
#| fig-width: 10
#| fig-height: 8

affiliation_data <- authors_data |>
  filter(!is.na(author_affiliation), author_affiliation != "")

if (nrow(affiliation_data) > 0) {
  # Prepare data
  aff_count <- affiliation_data |>
    count(author_affiliation, sort = TRUE) |>
    slice_head(n = 15) |>
    mutate(author_affiliation = fct_reorder(author_affiliation, n))

  # Recycle colorblind palette colors for all affiliations
  n_aff <- nrow(aff_count)
  colors <- rep(colorblind_pal()(8), length.out = n_aff)

  # Create plot
  aff_count |>
    ggplot(aes(x = n, y = author_affiliation, fill = author_affiliation)) +
    geom_col(show.legend = FALSE) +
    labs(title = "Top 15 Institutional Affiliations",
         x = "Number of Author Entries",
         y = "Affiliation") +
    scale_fill_manual(values = colors)
} else {
  cat("No affiliation data available")
}
```

## Resource Type Characteristics

```{r size_by_type_table}
#| echo: false

zenodo_data |>
  filter(!is.na(resource_type)) |>
  group_by(resource_type) |>
  summarise(
    count = n(),
    avg_size_mb = round(mean(total_size_mb, na.rm = TRUE), 2),
    median_size_mb = round(median(total_size_mb, na.rm = TRUE), 2),
    max_size_mb = round(max(total_size_mb, na.rm = TRUE), 2)
  ) |>
  arrange(desc(count)) |>
  rename(
    `Resource Type` = resource_type,
    `Count` = count,
    `Avg Size (MB)` = avg_size_mb,
    `Median Size (MB)` = median_size_mb,
    `Max Size (MB)` = max_size_mb
  ) |>
  knitr::kable(caption = "File Size Statistics by Resource Type")
```

## GitHub Integration

```{r github_stats_calc}
#| echo: false
#| include: false

if (nrow(github_data) > 0) {
  github_stats <- zenodo_data |>
    mutate(has_github = record_id %in% github_data$record_id) |>
    summarise(
      total = n(),
      with_github = sum(has_github),
      pct_github = round(with_github / total * 100, 1)
    )
}
```

```{r github_coverage_text}
#| echo: false
#| results: asis

if (nrow(github_data) > 0) {
  cat("**GitHub Repository Links:**\n\n")
  cat("- Records with GitHub:", github_stats$with_github, "/", github_stats$total,
      "(", github_stats$pct_github, "%)\n\n")
} else {
  cat("*No GitHub data available*\n\n")
}
```

```{r github_relation_table}
#| echo: false

if (nrow(github_data) > 0) {
  github_data |>
    count(github_relation, sort = TRUE) |>
    rename(Relation = github_relation, Count = n) |>
    knitr::kable(caption = "GitHub Relation Types")
}
```

```{r github_by_type}
#| fig-width: 10
#| fig-height: 6

if (nrow(github_data) > 0) {
  # Prepare data
  github_by_type_data <- zenodo_data |>
    mutate(has_github = record_id %in% github_data$record_id) |>
    filter(!is.na(resource_type)) |>
    count(resource_type, has_github)

  # Create plot
  github_by_type_data |>
    ggplot(aes(x = resource_type, y = n, fill = has_github)) +
    geom_col(position = "stack") +
    labs(title = "GitHub Integration by Resource Type",
         x = "Resource Type",
         y = "Number of Publications",
         fill = "Has GitHub Link") +
    scale_fill_colorblind(labels = c("No", "Yes")) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
} else {
  cat("No GitHub data available for visualization")
}
```

**Note:**

- "issupplementto": Zenodo record supplements the GitHub repository (e.g., data/manuscripts)
- "compiles": Zenodo record is compiled from the GitHub repository (e.g., software releases)

## Summary Statistics

```{r summary_stats}
#| echo: false

zenodo_data |>
  summarise(
    total_records = n(),
    total_size_gb = sum(total_size_mb, na.rm = TRUE) / 1024,
    avg_size_mb = mean(total_size_mb, na.rm = TRUE),
    median_size_mb = median(total_size_mb, na.rm = TRUE),
    avg_authors = mean((authors_data |> count(record_id))$n, na.rm = TRUE)
  ) |>
  knitr::kable(digits = 2, caption = "Summary Statistics")
```
